### 2020-01-23 presentation_templates/LaTex_beamer/slides.tex
Add presentation templates

\documentclass[xcolor=x11names,compress,usenames,dvipsnames,mathsans]{beamer}
\usepackage{textpos}
\usepackage[greek,english]{babel}
\usepackage{natbib}
\usepackage{wrapfig,color,dashrule,epsfig,graphicx}
\usepackage{pstool,psfrag}
\newcommand{\calt}{\color{red} } % alert color
\newcommand{\ccom}{\color{blue} } % commentary color

\newcommand{\redtext}[1]{\color{UOEred}{#1}\color{UOEblue}}

\newcommand{\link}[1]{\textbf{\textcolor{PMS3025}{\underline{#1}}}}
\definecolor{NewRed}{rgb}{0.8, 0.0, 0.0}

 %% Beamer Layout %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usetheme{Edinburgh}
\usecolortheme{edinburgh}
\usefonttheme[onlylarge]{structurebold}
\setbeamertemplate{itemize items}[circle]

\setbeamerfont{title}{size={\fontsize{13}{20}}}
\setbeamercolor*{palette tertiary}{fg=black,bg=black!10} 
\setbeamercolor*{palette quaternary}{fg=black,bg=black!10} 
\setbeamercolor{alerted text}{fg=red}
\setbeamercolor{sidebar}{bg=red!25!blue,fg=white}

\def\ci{\perp\!\!\!\perp}

\setbeamercovered{transparent}

\addtobeamertemplate{footline}{}
{
  \begin{textblock*}{100mm}(.845\textwidth,-0.57cm)

    \includegraphics[height=0.5cm, width=1.8cm]{225px-Epcc_logo.jpg}
  \end{textblock*}
}

\addtobeamertemplate{frametitle}{}
{
  \begin{textblock*}{100mm}(.935\textwidth,-0.645cm) 
	\includegraphics[scale=.1]{crest_rb.eps}
  \end{textblock*}
}

\usepackage{graphicx} % Allows including images
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables

%\setbeamercovered{transparent}

%----------------------------------------------------------------------------------------
%	STYLE STUFF -- section title slides
%----------------------------------------------------------------------------------------

\AtBeginSection[]{
  \begin{frame}
  \vfill
  \centering
  \begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
    \usebeamerfont{title}\insertsectionhead\par%
  \end{beamercolorbox}
  \vfill
  \end{frame}
}

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[GitLab demo]{Project Preparation \\ GitLab} 
% The [Short Title] appears at the bottom of every slide, the {Full Title} is only on the title page

\author[author]{Full Name \\
\textit{email@ed.ac.uk} \\
\vspace{1em}
\includegraphics[scale=.8]{logo_colour.pdf}
} % Your name

\date{\today}

\begin{document}

\begin{frame}
\titlepage % Print the title page as the first slide

\end{frame}

\begin{frame}
\frametitle{Overview} % Table of contents slide, comment this block out to remove it
\tableofcontents % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
\end{frame}

%----------------------------------------------------------------------------------------
%	PRESENTATION SLIDES
%----------------------------------------------------------------------------------------

%------------------------------------------------
%\section{Section Title} % Sections can be created in order to organize your presentation into discrete blocks, all sections and subsections are automatically printed in the table of contents as an overview of the talk
%------------------------------------------------

\begin{frame}
\frametitle{Purpose}
\vfill
GitLab is a version control service provided by the University of Edinburgh. \\
\vfill
\pause
It's purpose in the MSc project work is to:
\begin{itemize}
	\item maintain traceable history
	\item document your progress
	\item share information with your supervisor(s) and markers
\end{itemize}
\pause
\vfill
(and it will be marked accordingly.)
\vfill
\end{frame}

%------------------------------------------------
\begin{frame}
\frametitle{Content}

%Proof of your engagement with the project work and ability to organise your work and materials.
%\pause

Project Information:
	\begin{itemize}
		\item Project Title
		\item Student's Name and UUN
		\item EPCC Supervisor's Name
		\item (External Supervisor's Name, Email and Organisation)
	\end{itemize}

\vfill
\pause

Project Materials:
	\begin{itemize}
		\item initial and final project proposal
		\item meeting minutes
		\item Project Preparation report + slides
		\item Dissertation report + slides
		\item code/scripts
	\end{itemize}

\end{frame}

%------------------------------------------------

\begin{frame}
\frametitle{Further Information}

\begin{itemize}
	\item \href{https://git.ecdf.ed.ac.uk/}{\link{Log in page}}
	\item \href{https://www.wiki.ed.ac.uk/pages/viewpage.action?spaceKey=ResearchServices\&title=Version+Control+Service+-+GitLab}{\link{GitLab wiki}}
\end{itemize}

\end{frame}

%------------------------------------------------

\end{document}

### 2020-02-17 presentation/slides.tex
Add presentation

\documentclass[xcolor=x11names,compress,usenames,dvipsnames,mathsans]{beamer}
\usepackage{textpos}
\usepackage[greek,english]{babel}
\usepackage{natbib}
\usepackage{wrapfig,color,dashrule,epsfig,graphicx}
\usepackage{pstool,psfrag}
\newcommand{\calt}{\color{red} } % alert color
\newcommand{\ccom}{\color{blue} } % commentary color

\newcommand{\redtext}[1]{\color{UOEred}{#1}\color{UOEblue}}

\newcommand{\link}[1]{\textbf{\textcolor{PMS3025}{\underline{#1}}}}
\definecolor{NewRed}{rgb}{0.8, 0.0, 0.0}

 %% Beamer Layout %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usetheme{Edinburgh}
\usecolortheme{edinburgh}
\usefonttheme[onlylarge]{structurebold}
\setbeamertemplate{itemize items}[circle]

\setbeamerfont{title}{size={\fontsize{13}{20}}}
\setbeamercolor*{palette tertiary}{fg=black,bg=black!10} 
\setbeamercolor*{palette quaternary}{fg=black,bg=black!10} 
\setbeamercolor{alerted text}{fg=red}
\setbeamercolor{sidebar}{bg=red!25!blue,fg=white}

\def\ci{\perp\!\!\!\perp}

\setbeamercovered{transparent}

\addtobeamertemplate{footline}{}
{
  \begin{textblock*}{100mm}(.845\textwidth,-0.57cm)

    \includegraphics[height=0.5cm, width=1.8cm]{225px-Epcc_logo.jpg}
  \end{textblock*}
}

\addtobeamertemplate{frametitle}{}
{
  \begin{textblock*}{100mm}(.935\textwidth,-0.645cm) 
	\includegraphics[scale=.1]{crest_rb.eps}
  \end{textblock*}
}

\usepackage{graphicx} % Allows including images
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables

%\setbeamercovered{transparent}

%----------------------------------------------------------------------------------------
%	STYLE STUFF -- section title slides
%----------------------------------------------------------------------------------------

\AtBeginSection[]{
  \begin{frame}
  \vfill
  \centering
  \begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
    \usebeamerfont{title}\insertsectionhead\par%
  \end{beamercolorbox}
  \vfill
  \end{frame}
}

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[Parallel-in-time methods with ML]{Parallel-in-time methods with ML} 
% The [Short Title] appears at the bottom of every slide, the {Full Title} is only on the title page

\author[author]{Viktor Csomor \\
\textit{s1984842@ed.ac.uk} \\
\vspace{1em}
\includegraphics[scale=.8]{logo_colour.pdf}
} % Your name

\date{\today}

\begin{document}

\begin{frame}
\titlepage % Print the title page as the first slide
\begin{center}
Supervisors:
Rupert Nash, Anna Roubíčková
\end{center}
\end{frame}

\begin{frame}
\frametitle{Overview} % Table of contents slide, comment this block out to remove it
\tableofcontents % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
\end{frame}

%----------------------------------------------------------------------------------------
%	PRESENTATION SLIDES
%----------------------------------------------------------------------------------------

%------------------------------------------------
%\section{Section Title} % Sections can be created in order to organize your presentation into discrete blocks, all sections and subsections are automatically printed in the table of contents as an overview of the talk
%------------------------------------------------

\section{Motivation}

\begin{frame}
\frametitle{Differential equations}
\begin{itemize}
	\item numerous scientific and engineering problems are described by time-dependent differential equations
	\item ODEs and PDEs
	\item many different numerical methods exist for solving them \cite{suli2003}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Parallelism}
\begin{itemize}
	\item traditionally, parallelised across the system or across space \cite{gear1988}
	\item good weak scaling, limited strong scaling
	\item but for some problems only wall time matters (e.g. stock trading)
	\item need to exploit parallelism better to reduce wall clock time
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Parallel-in-time methods}
\begin{itemize}
	\item solution parallelised across the time steps
	\item more CPU time, less wall clock time
	\item parareal algorithm \cite{parareal} for solving IVPs
	\item relies on two operators, coarse $G$ and fine $F$
	\item for every iteration, $G$ executed serially, $F$ in parallel
	\item $G$ must be fast but preferably also accurate enough
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Parareal algorithm}
\begin{figure}[!htb]
\begin{center}
\resizebox{1.0\hsize}{!}{\includegraphics{parareal.jpg}}
\end{center}
\end{figure}
\end{frame}

%------------------------------------------------

\section{Hypotheses}

\begin{frame}
\frametitle{Hypotheses}
\begin{enumerate}
	\item an ML (regression) model can be trained and used as $G$ to solve differential equations in a parareal framework
	\item for some problems, the ML model based $G$ operator is better than traditional coarse operators in terms of performance and accuracy
	\item for some problems, the ML accelerated parareal framework achieves shorter wall time than a conventional parareal algorithm or a state-of-the-art traditional solver using the same number of CPU cores
	\begin{enumerate}[a]
	    \item training time ignored
	    \item training time taken into account
	\end{enumerate}
\end{enumerate}
\end{frame}

%------------------------------------------------

\section{Related work}

\begin{frame}
\frametitle{Related work}
\begin{itemize}
    \item neural network for solving time-dependent differential equations \cite{regazzoni2019}
    \item neural network to approximate the gradients of high-dimensional PDEs \cite{han2018}
    \item physics informed neural network (PINN) for solving PDEs \cite{raissi2017} \cite{sirignano2018} \cite{lu2019}
    \item parareal solver using a PINN as $F$ \cite{meng2019}
\end{itemize}
\end{frame}

%------------------------------------------------

\section{Challenges}

\begin{frame}
\frametitle{Challenges}
\begin{itemize}
    \item numerical analysis is a complex field
    \item fast moving area of research
    \item many potential avenues (PDE solving methods, model training, tech stack)
    \item aiming high
\end{itemize}
\end{frame}

%------------------------------------------------

\section{Plan}

\begin{frame}
\frametitle{Implementation}
\begin{itemize}
    \item develop a simple parareal framework
    \begin{itemize}
        \item[$-$] already implemented first version in C with POSIX threads 
    \end{itemize}
    \item construct an ML model (PINN vs standard models)
    \item train the model using targets provided either by $G$ or $F$
    \item automatise the model training as part of the framework
    \item use the trained model for inference as $G$
    \item extend the framework to discretise PDEs or handle PDEs directly through the model
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Problem iterations}
\begin{itemize}
    \item ODE (Lotka-Volterra)
    \item simple PDE (2D diffusion)
    \item high-dimensional PDE (Black-Scholes)
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Performance}
\begin{itemize}
    \item compare the ML version of $G$ to traditional ODE solvers of similar accuracy
    \item compare the performance of the parareal framework to traditional ODE and PDE solvers and analyse the impact of the model training on performance
    \item try different ML models (linear regression, ANN, RNN, etc.)
    \item take advantage of Cirrus' new GPUs
\end{itemize} 
\end{frame}

\begin{frame}
\frametitle{Tech stack}
\begin{itemize}
    \item C/C++ faster than Python but most scientific Python libraries are just wrappers on top of C bindings
    \item Python better suited for ML and numerical solver libraries are easier to use
    \item mpi4py \cite{dalcin2005} for distributed computing
    \item scikit-learn \cite{pedregosa2011} and keras \cite{chollet2015} for ML
    \item FEniCS \cite{fenics2015} (finite element) and FiPy \cite{fipy2009} (finite volume) to compare performance against
\end{itemize}
\end{frame}

%------------------------------------------------

\begin{frame}[allowframebreaks]
\frametitle{References}
\bibliographystyle{unsrt}
\bibliography{ref}
\end{frame}

\end{document}

### 2020-03-22 report/MSc-report.tex
Add report template

\documentclass{article}

\usepackage{epcc}
\usepackage{graphics}
\usepackage{booktabs}

\renewcommand{\labelenumii}{\theenumii}
\renewcommand{\theenumii}{\theenumi.\arabic{enumii}.}
% this will make nested numbering all arabic numbers

% This example file shows how a report can be laid out using Latex. It
% does not use any special local features so should be portable to other
% places.
% 
% When producing draft copies of a thesis you may want to print only
% selected pages of the thesis. To do this use the command
% 
% dvips -f -p 4 -n 3 myfile.dvi | lpr
% 
% where -p 4 means start printing at page 4 (ie the page that will be
% numbered 4, not necessarily the 4th page) and -n 3 means print 3 pages.
% This example will print pages 4, 5 and 6.
% 
% If you want to print the report and also save paper you can print more
% than one page on each sheet of paper. Use the command
% 
% dvips -f myfile.dvi | psnup -2 | lpr
% 
% to print 2 pages per sheet. psnup can take values 2, 4, 8, or 9.
%
% To produce a PDF version you can create a PostScript copy first
%
% dvips -f myfile.dvi > myfile.pdf
%
% and then convert it
%
% distill myfile.ps
%
% or you can go straight to PDF
%
% pdflatex myfile
%
% Note that pdflatex expects all included figures to be in PDF too. See
% the includegraphics command below.

% This document contains many cross-references and forward references,
% eg in constructing a table of contents, so Latex may need to be run
% twice to get all the references correct. If you need to run Latex twice
% you may get the warning:
% 
% LaTeX Warning: Label(s) may have changed. Rerun to get cross-references right.

\begin{document}

\pagenumbering{roman}

\title{A \LaTeX{}  Report Example}
\author{David Henty \& MSc Dissertations Team}
\date{\today}

\makeEPCCtitle

\newpage

\tableofcontents

\newpage

\section*{Acknowledgements}

This template is a slightly modified version of the one developed by
Prof. Charles Duncan for MSc students in the Dept. of Meteorology. His
acknowledgement follows:

{\em This template has been produced with help from many former students who
have shown different ways of doing things. Please make suggestions for
further improvements.}

\section*{Note}

This template was originally created for a full MSc dissertation.
Although I have modified the format slightly (eg there are no separate
chapters), I have not substantially changed the text or the advice
appearing in the \LaTeX comments (source lines beginning with \%). You
should therefore be aware that for a shorter report you may
not need to go into as much detail as indicated here. For example, you
might not need to include any acknowledgements and any reference list is
likely to be very short.

\newpage
\pagenumbering{arabic}

\section{Introduction}

This should contain a description of your project and the problem you
are trying to solve. Where appropriate you should also include
references to work which has already been done on your topic and
anything else which lets you set your work in context.

One of the things you will need to do is to ensure that you have a
suitable list of references.  To do this you should see \cite{ref:lam}
or some other suitable reference.  Note the format of the citation used
here is the style favoured in this department.  Here is another
reference \cite{ref:bloggs} for good measure.

You will also want to make sure you have no spelling or grammatical
mistakes. To help idwentify spelling mistukes you caan use the commands
{\em ispell} or {\em spell}. See the appropriate manual pages. Remember
that spelling mistakes are not the only errors which can occur. Spelling
checkers will not find errors which are, in fact, valid words such as
{\em there} for {\em their}, nor will they find repeated words which
sometimes occur if your concentration is broken when typing. {\bf There
is no substitute for thorough proof reading!}

\section{Previous Dissertation Review}

In the introductory paragraph, clearly identify the reviewed dissertation by its Title, Author and Year.

\subsection*{Summarize the project in your own words.}

This is just to show how to break things into sections.

Here is a padding paragraph.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More
rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.   Rhubarb.  More
rhubarb.  Yet more rhubarb.   Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.    Rhubarb.  More
rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.

\subsection*{Does the dissertation adequately explain the context in which the problem is set?}

Give evidence to support your answer based on the quality of motivation, background and previous work overview.

Here is a padding paragraph.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More
rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.   Rhubarb.  More
rhubarb.  Yet more rhubarb.   Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.    Rhubarb.  More
rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.

\subsection*{Has the student adequately explained the scope of the problem, the approach that they would take, how success would be measured and reflected on the outcomes?}

Consider limitations, reproducibility and research methods used. Give evidence to support your answer.

\subsection*{Quality of presentation.}
Comment on the overall structure, whether the explanations were clear and easy to follow, how was the language used.

\subsection*{Presentation of additional information.}
Discus the use of references and citations throughout the text and use of figures and tables (their design, placement and descriptions, references to from within text, etc.)

\section{Background and Literature Review}

I created a figure which is centred and stretched to 30\% of the
width of the page \verb+{0.30\hsize}+ and with the height stretched by
the same amount \verb+{!}+ to preserve the aspect ratio. If you omit the
extension (ie .eps, .ps or .pdf) on the file name then \LaTeX  will pick
up the postscript copy whereas pdflatex will automatically pick up the
PDF version.

\begin{figure}

\begin{center}
\resizebox{0.30\hsize}{!}{\includegraphics{logos/crest_bw}}
\end{center}

\caption{The University Crest}
\label{fig:eucrest}

\end{figure}

% see the man page for dvips for details of the special command which is
% much more powerful than is shown here. It allows offsets in the
% horizontal and vertical and scaling in x and y.

% choosing suitable values for offset and scale can be a tiring matter
% of trial and error.

% note that labels do not need to include a description of the object
% they are labelling but it can be helpful, eg \label{fig:figurename}.

You can use a label on a figure to refer to it later: The university
crest is in Figure~\ref{fig:eucrest}. Note that you should not use phrases like
``the figure above'' or ``the following figure'' since \LaTeX  may move
the figure relative to the text if it cannot be fitted onto the current page.

\section{Preliminary Investigations}

You might want to include an equation here:

\begin{equation} \delta N_{\nu} = (\delta N_{\nu})_{ex} + (\delta N_{\nu})_{au} 
\label{equation:delsplit}
% note that the label is optional but allows you to refer to this
% equation later.
\end{equation}

You might sometimes want to include equations without numbering them.

\begin{displaymath}
E=mc^{2}
\end{displaymath}

You might also want to include diagrams.  The example shows the use of
the special command which allows existing postscript files to be
included.  You would normally keep your figures separate from the text. 
These pictures might be satellite images or postscript output from some
program such as IDL, PV-WAVE, Uniras or xpaint.

\paragraph{More on numbering:}

This text is in a paragraph which is also not numbered by default and
the ``title'' of the paragraph is not on a separate line.
If you want to increase the depth to which subsections are numbered you
should see the subsection on setting the secnumdepth counter in the manual.

\section{Final Proposal}

One way to produce simple graphs is to use gnuplot which can produce
native \LaTeX  output. Graph \ref{fig:gnu} was produced using gnuplot with
output designated as \LaTeX; this creates a \LaTeX output file which you
can include directly or keep separate and refer to using the {\em
include} command.

Another approach is to draw simple figures using {\em xfig} which allows
you to export diagrams in \LaTeX picture format so that the diagram can
be included directly.

Perhaps the most robust way to include graphs is to create then using
whatever package you like (eg Excel) and then convert them to PostScript
or PDF and include them in the same was as was done in
Figure~\ref{fig:eucrest} for the University Crest. You can usually do
this with most packages, including Microsoft ones. One trick for
producing PostScript is to print to a dummy PostScript printer; for PDF
you can use the free CutePDF utilities at {\verb+www.cutepdf.com+}.

% in practice you would probably keep this in a separate file and use
% the \include{filename} command to insert it here.

\begin{figure}
% GNUPLOT: LaTeX picture
\setlength{\unitlength}{0.240900pt}
\ifx\plotpoint\undefined\newsavebox{\plotpoint}\fi
\sbox{\plotpoint}{\rule[-0.200pt]{0.400pt}{0.400pt}}%
\begin{picture}(1500,1200)(0,0)
\font\gnuplot=cmr10 at 10pt
\gnuplot
\sbox{\plotpoint}{\rule[-0.200pt]{0.400pt}{0.400pt}}%
\put(220.0,113.0){\rule[-0.200pt]{292.934pt}{0.400pt}}
\put(220.0,113.0){\rule[-0.200pt]{0.400pt}{245.477pt}}
\put(220.0,113.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
\put(198,113){\makebox(0,0)[r]{$0$}}
\put(1416.0,113.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
\put(220.0,317.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
\put(198,317){\makebox(0,0)[r]{$0.2$}}
\put(1416.0,317.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
\put(220.0,521.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
\put(198,521){\makebox(0,0)[r]{$0.4$}}
\put(1416.0,521.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
\put(220.0,724.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
\put(198,724){\makebox(0,0)[r]{$0.6$}}
\put(1416.0,724.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
\put(220.0,928.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
\put(198,928){\makebox(0,0)[r]{$0.8$}}
\put(1416.0,928.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
\put(220.0,1132.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
\put(198,1132){\makebox(0,0)[r]{$1$}}
\put(1416.0,1132.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
\put(220.0,113.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
\put(220,68){\makebox(0,0){$0$}}
\put(220.0,1112.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
\put(414.0,113.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
\put(414,68){\makebox(0,0){$1$}}
\put(414.0,1112.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
\put(607.0,113.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
\put(607,68){\makebox(0,0){$2$}}
\put(607.0,1112.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
\put(801.0,113.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
\put(801,68){\makebox(0,0){$3$}}
\put(801.0,1112.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
\put(995.0,113.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
\put(995,68){\makebox(0,0){$4$}}
\put(995.0,1112.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
\put(1188.0,113.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
\put(1188,68){\makebox(0,0){$5$}}
\put(1188.0,1112.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
\put(1382.0,113.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
\put(1382,68){\makebox(0,0){$6$}}
\put(1382.0,1112.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
\put(220.0,113.0){\rule[-0.200pt]{292.934pt}{0.400pt}}
\put(1436.0,113.0){\rule[-0.200pt]{0.400pt}{245.477pt}}
\put(220.0,1132.0){\rule[-0.200pt]{292.934pt}{0.400pt}}
\put(45,622){\makebox(0,0){\shortstack{This is\\the\\$y$ axis}}}
\put(828,23){\makebox(0,0){This is the $x$ axis}}
\put(828,1177){\makebox(0,0){This is a plot of $y=sin(x)$}}
\put(220.0,113.0){\rule[-0.200pt]{0.400pt}{245.477pt}}
\sbox{\plotpoint}{\rule[-0.500pt]{1.000pt}{1.000pt}}%
\put(1306,1067){\makebox(0,0)[r]{sin(x)}}
\multiput(1328,1067)(20.756,0.000){4}{\usebox{\plotpoint}}
\put(1394,1067){\usebox{\plotpoint}}
\put(220,113){\usebox{\plotpoint}}
\multiput(220,113)(3.768,20.411){4}{\usebox{\plotpoint}}
\multiput(232,178)(4.132,20.340){3}{\usebox{\plotpoint}}
\multiput(245,242)(3.825,20.400){3}{\usebox{\plotpoint}}
\multiput(257,306)(3.884,20.389){3}{\usebox{\plotpoint}}
\multiput(269,369)(3.944,20.377){3}{\usebox{\plotpoint}}
\multiput(281,431)(4.326,20.300){3}{\usebox{\plotpoint}}
\multiput(294,492)(4.137,20.339){3}{\usebox{\plotpoint}}
\multiput(306,551)(4.276,20.310){3}{\usebox{\plotpoint}}
\multiput(318,608)(4.693,20.218){3}{\usebox{\plotpoint}}
\multiput(331,664)(4.583,20.243){2}{\usebox{\plotpoint}}
\multiput(343,717)(4.754,20.204){3}{\usebox{\plotpoint}}
\multiput(355,768)(5.034,20.136){2}{\usebox{\plotpoint}}
\multiput(367,816)(5.760,19.940){2}{\usebox{\plotpoint}}
\multiput(380,861)(5.579,19.992){3}{\usebox{\plotpoint}}
\put(398.00,923.50){\usebox{\plotpoint}}
\multiput(404,943)(7.049,19.522){2}{\usebox{\plotpoint}}
\multiput(417,979)(7.288,19.434){2}{\usebox{\plotpoint}}
\put(433.18,1021.10){\usebox{\plotpoint}}
\multiput(441,1040)(8.982,18.712){2}{\usebox{\plotpoint}}
\put(460.41,1076.97){\usebox{\plotpoint}}
\put(471.84,1094.28){\usebox{\plotpoint}}
\put(484.84,1110.41){\usebox{\plotpoint}}
\put(500.42,1124.01){\usebox{\plotpoint}}
\multiput(503,1126)(19.159,7.983){0}{\usebox{\plotpoint}}
\put(519.48,1131.37){\usebox{\plotpoint}}
\multiput(527,1132)(20.136,-5.034){0}{\usebox{\plotpoint}}
\put(539.74,1128.60){\usebox{\plotpoint}}
\put(557.04,1117.38){\usebox{\plotpoint}}
\put(570.79,1101.95){\usebox{\plotpoint}}
\put(582.44,1084.80){\usebox{\plotpoint}}
\put(593.09,1066.99){\usebox{\plotpoint}}
\multiput(601,1053)(8.430,-18.967){2}{\usebox{\plotpoint}}
\put(619.18,1010.54){\usebox{\plotpoint}}
\multiput(625,996)(7.413,-19.387){2}{\usebox{\plotpoint}}
\multiput(638,962)(6.403,-19.743){2}{\usebox{\plotpoint}}
\multiput(650,925)(5.830,-19.920){2}{\usebox{\plotpoint}}
\multiput(662,884)(5.461,-20.024){2}{\usebox{\plotpoint}}
\multiput(674,840)(5.533,-20.004){2}{\usebox{\plotpoint}}
\multiput(687,793)(4.937,-20.160){3}{\usebox{\plotpoint}}
\multiput(699,744)(4.667,-20.224){2}{\usebox{\plotpoint}}
\multiput(711,692)(4.858,-20.179){3}{\usebox{\plotpoint}}
\multiput(724,638)(4.276,-20.310){3}{\usebox{\plotpoint}}
\multiput(736,581)(4.205,-20.325){3}{\usebox{\plotpoint}}
\multiput(748,523)(4.070,-20.352){3}{\usebox{\plotpoint}}
\multiput(760,463)(4.326,-20.300){3}{\usebox{\plotpoint}}
\multiput(773,402)(3.884,-20.389){3}{\usebox{\plotpoint}}
\multiput(785,339)(3.884,-20.389){3}{\usebox{\plotpoint}}
\multiput(797,276)(4.070,-20.352){3}{\usebox{\plotpoint}}
\multiput(810,211)(3.825,-20.400){3}{\usebox{\plotpoint}}
\multiput(822,147)(3.607,-20.440){2}{\usebox{\plotpoint}}
\put(828,113){\usebox{\plotpoint}}
\end{picture}
\caption{Simple Gnuplot example}
\label{fig:gnu}
\end{figure}

\section{Workplan}

You are likely to use tables and graphs. You can create tables easily in \LaTeX.

\begin{table}[h]
\begin{center}
\begin{tabular}{ l|c|l }
\toprule
{\bf File names} & {\bf Satellite} & {\bf Resolution}\\
\midrule
  worldr         &  Meteosat     &   5km\\
  worldg         &  Meteosat     &   5km\\
  worldb         &  Meteosat     &   5km\\
\bottomrule
\end{tabular}
\end{center}
\caption{This is a simple table. More complicated tables can have
headings which pass over more than one column}
\label{simple_table}
\end{table}

If you want to produce fancier tables than shown in Table \ref{simple_table}
refer to the manual.

\section{Risk Analysis}

Here is a padding paragraph.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More
rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.

\section{Outline of the Dissertation Report}

You will probably need a numbered list here, which is rather simple in \LaTeX:

\begin{enumerate}
	\item This is top level rhubarb.
	\begin{enumerate}
		\item 2nd level
		\item some rhubarb
		\item and more
	\end{enumerate}
	\item More top level rhubarb
	\begin{enumerate}
		\item other boring vegetables 
		\item kale, for example
	\end{enumerate}
	\item And so on
\end{enumerate}

\pagebreak
\begin{thebibliography}{100}

\bibitem{ref:lam} L.Lamport. {\em 1986 Latex User's Guide
and Reference Manual.} Addison Wesley. pp242.

\bibitem{ref:bloggs} F.Bloggs. {\em 1993 Latex Users do it
in Environments} Int. Journal of Silly Findings. pp 23-29.

\end{thebibliography}

\end{document}

### 2020-03-22 code/python/src/runge_kutta.py
Update documentation

@@ -3,7 +3,7 @@ from typing import Callable
 
 class ForwardEulerMethod:
     """
-    The forward Euler method, an explicit first order Runge-Kutta procedure.
+    The forward Euler method, an explicit first order Runge-Kutta method.
     """
 
     def __call__(self, y: float, x: float, d_x: float, d_y_wrt_x: Callable[[float, float], float]) -> float:
@@ -12,7 +12,7 @@ class ForwardEulerMethod:
 
 class ExplicitMidpointMethod:
     """
-    The explicit midpoint method, a second order Runge-Kutta procedure.
+    The explicit midpoint method, a second order Runge-Kutta method.
     """
 
     def __call__(self, y: float, x: float, d_x: float, d_y_wrt_x: Callable[[float, float], float]) -> float:
@@ -21,7 +21,7 @@ class ExplicitMidpointMethod:
 
 class RK4:
     """
-    The RK4 method, an explicit fourth order Runge-Kutta procedure.
+    The RK4 method, an explicit fourth order Runge-Kutta method.
     """
 
     def __call__(self, y: float, x: float, d_x: float, d_y_wrt_x: Callable[[float, float], float]) -> float:

### 2020-03-22 code/python/src/parareal.py
Add Parareal implementation using Python threading

@@ -1,54 +1,87 @@
 import numpy as np
 import math
 
-from mpi4py import MPI
+import threading
 
 from src.runge_kutta import ForwardEulerMethod, RK4
 
 r = .01
-n0 = 10000.
+n_0 = 10000.
+
+fine_d_t = .25
+coarse_d_t = 2 * fine_d_t
+k = 2
+
+duration = 8.
+n_threads = 8
+time_slices = np.arange(0., duration, duration / n_threads)
 
 g = ForwardEulerMethod()
 f = RK4()
-coarse_d_t = .5
-fine_d_t = .25
-k = 2
 
-duration = .5
-time_steps = int(duration / coarse_d_t)
+
+def operator(integration_method, y_0, t_0, t_max, d_t, d_y_wrt_t):
+    y = [y_0]
+    y_i = y_0
+    for i, t in enumerate(np.arange(t_0, t_max, d_t)):
+        t_i = t_0 + i * d_t
+        y_i = integration_method(y_i, t_i, d_t, d_y_wrt_t)
+        y.append(y_i)
+    return y[-1]
+
+
+def calculate_corrections(y_0, t_0, t_max, d_y_wrt_t, slice_ind, corr):
+    f_value = operator(f, y_0, t_0, t_max, fine_d_t, d_y_wrt_t)
+    g_value = operator(g, y_0, t_0, t_max, coarse_d_t, d_y_wrt_t)
+    corr[slice_ind] = f_value - g_value
 
 
 def d_rabbit_population_wrt_t(_, n): return r * n
 
 
-comm = MPI.COMM_WORLD
-rank = comm.Get_rank()
-size = comm.Get_size()
+n = np.empty(len(time_slices))
+n_corr = np.empty(len(time_slices) - 1)
+n_exact = np.empty(len(time_slices))
 
-n = np.empty(time_steps)
-n_corr = np.empty(time_steps)
-n_exact = np.empty(time_steps)
+n[0] = n_0
+n_exact[0] = n_0
 
-n[0] = n0
-n_corr[0] = n0
-n_exact[0] = n0
+for i, t in enumerate(time_slices[1:]):
+    n_exact[i + 1] = n_0 * math.exp(r * t)
 
-for i in range(0, time_steps - 1):
-    t = i * coarse_d_t
-    n[i + 1] = g(n[i], t, coarse_d_t, d_rabbit_population_wrt_t)
-
-    t_next = t + coarse_d_t
[…]

### 2020-03-23 code/python/src/parareal.py
Implement parallelism using mpi4py

@@ -1,9 +1,9 @@
 import numpy as np
 import math
 
-import threading
+from mpi4py import MPI
 
-from src.runge_kutta import ForwardEulerMethod, RK4
+from runge_kutta import ForwardEulerMethod, RK4
 
 r = .01
 n_0 = 10000.
@@ -12,12 +12,15 @@ fine_d_t = .25
 coarse_d_t = 2 * fine_d_t
 k = 2
 
-duration = 8.
-n_threads = 8
-time_slices = np.arange(0., duration, duration / n_threads)
-
 g = ForwardEulerMethod()
 f = RK4()
+
+comm = MPI.COMM_WORLD
+rank = comm.Get_rank()
+size = comm.Get_size()
+
+duration = 8.
+time_slices = np.linspace(0., duration, size + 1)
 
 
 def operator(integration_method, y_0, t_0, t_max, d_t, d_y_wrt_t):
@@ -30,10 +33,10 @@ def operator(integration_method, y_0, t_0, t_max, d_t, d_y_wrt_t):
     return y[-1]
 
 
-def calculate_corrections(y_0, t_0, t_max, d_y_wrt_t, slice_ind, corr):
+def calculate_corrections(y_0, t_0, t_max, d_y_wrt_t):
     f_value = operator(f, y_0, t_0, t_max, fine_d_t, d_y_wrt_t)
     g_value = operator(g, y_0, t_0, t_max, coarse_d_t, d_y_wrt_t)
-    corr[slice_ind] = f_value - g_value
+    return f_value - g_value
 
 
 def d_rabbit_population_wrt_t(_, n): return r * n
@@ -61,16 +64,8 @@ for i, t in enumerate(time_slices[:-1]):
 n_coarse = np.copy(n)
 
 for i in range(k):
-    threads = []
-    for j, t in enumerate(time_slices[:-1]):
-        thread = threading.Thread(
-            target=calculate_corrections,
-            args=(n[j], t, time_slices[j + 1], d_rabbit_population_wrt_t, j, n_corr))
-        thread.start()
-        threads.append(thread)
-
-    for j in range(len(threads)):
-        threads[j].join()
+    n_corr_rank = calculate_corrections(n[rank], time_slices[rank], time_slices[rank + 1], d_rabbit_population_wrt_t)
+    comm.Allgather([n_corr_rank, MPI.DOUBLE],  [n_corr, MPI.DOUBLE])
 
     for j, t in enumerate(time_slices[:-1]):
[…]

### 2020-03-23 code/python/src/runge_kutta.py
Add a base class for Runge-Kutta methods

@@ -1,30 +1,42 @@
 from typing import Callable
 
 
-class ForwardEulerMethod:
+class RungeKuttaMethod:
+    """
+    A base class for Runge-Kutta methods
+    """
+
+    """
+    Estimates the value of y at x^ = x + d_x.
+    """
+    def integrate(self, y: float, x: float, d_x: float, d_y_wrt_x: Callable[[float, float], float]) -> float:
+        pass
+
+
+class ForwardEulerMethod(RungeKuttaMethod):
     """
     The forward Euler method, an explicit first order Runge-Kutta method.
     """
 
-    def __call__(self, y: float, x: float, d_x: float, d_y_wrt_x: Callable[[float, float], float]) -> float:
+    def integrate(self, y: float, x: float, d_x: float, d_y_wrt_x: Callable[[float, float], float]) -> float:
         return y + d_x * d_y_wrt_x(x, y)
 
 
-class ExplicitMidpointMethod:
+class ExplicitMidpointMethod(RungeKuttaMethod):
     """
     The explicit midpoint method, a second order Runge-Kutta method.
     """
 
-    def __call__(self, y: float, x: float, d_x: float, d_y_wrt_x: Callable[[float, float], float]) -> float:
+    def integrate(self, y: float, x: float, d_x: float, d_y_wrt_x: Callable[[float, float], float]) -> float:
         return y + d_x * d_y_wrt_x(x + d_x / 2., y + d_y_wrt_x(x, y) * d_x / 2.)
 
 
-class RK4:
+class RK4(RungeKuttaMethod):
     """
     The RK4 method, an explicit fourth order Runge-Kutta method.
     """
 
-    def __call__(self, y: float, x: float, d_x: float, d_y_wrt_x: Callable[[float, float], float]) -> float:
+    def integrate(self, y: float, x: float, d_x: float, d_y_wrt_x: Callable[[float, float], float]) -> float:
         k1 = d_x * d_y_wrt_x(x, y)
         k2 = d_x * d_y_wrt_x(x + d_x / 2., y + k1 / 2.)
         k3 = d_x * d_y_wrt_x(x + d_x / 2., y + k2 / 2.)

### 2020-03-23 code/python/tests/test_runge_kutta.py

@@ -1,6 +1,6 @@
 import pytest
 
-from src.runge_kutta import ForwardEulerMethod, ExplicitMidpointMethod, RK4
+from runge_kutta import ForwardEulerMethod, ExplicitMidpointMethod, RK4
 
 
 def test_forward_euler_method():
@@ -13,7 +13,7 @@ def test_forward_euler_method():
     def d_y_wrt_x(x, y): return 5 * y + x ** 2
 
     expected_y2 = 4.
-    actual_y2 = euler(y1, x1, d_x, d_y_wrt_x)
+    actual_y2 = euler.integrate(y1, x1, d_x, d_y_wrt_x)
 
     assert actual_y2 == pytest.approx(expected_y2)
 
@@ -28,7 +28,7 @@ def test_explicit_midpoint_method():
     def d_y_wrt_x(x, y): return 2 * y - 4 * x
 
     expected_y2 = 4.5
-    actual_y2 = midpoint(y1, x1, d_x, d_y_wrt_x)
+    actual_y2 = midpoint.integrate(y1, x1, d_x, d_y_wrt_x)
 
     assert actual_y2 == pytest.approx(expected_y2)
 
@@ -43,6 +43,6 @@ def test_rk4():
     def d_y_wrt_x(x, y): return 2 * y + 1
 
     expected_y2 = 3.
-    actual_y2 = rk4(y1, x1, d_x, d_y_wrt_x)
+    actual_y2 = rk4.integrate(y1, x1, d_x, d_y_wrt_x)
 
     assert actual_y2 == pytest.approx(expected_y2)

### 2020-03-23 code/python/src/diff_eq.py
Create a base class for time dependent ODEs

@@ -1,16 +1,53 @@
+import math
 
 
-class DiffEq:
+class TimeDependentDiffEq:
     """
-    A representation of time dependent differential equations.
+    A representation of a time-dependent ordinary differential equation of the form y'(t) = f(t, y(t)).
     """
 
-    def __init__(self, independent_vars, conditions):
-        self.independent_vars = independent_vars
-        self.conditions = conditions
+    """
+    Returns whether the differential equation has an analytic solution
+    """
+    def has_exact_solution(self) -> bool:
+        pass
 
-    def get_independent_vars(self):
-        return self.independent_vars
+    """
+    Returns the exact value of y(t) given t.
+    """
+    def exact_y(self, t: float) -> float:
+        pass
 
-    def get_conditions(self):
-        return self.conditions
+    """
+    Returns the value of y(0).
+    """
+    def y_0(self) -> float:
+        pass
+
+    """
+    Returns the value of the derivative of y at t given t and y(t).
+    """
+    def d_y(self, t: float, y: float) -> float:
+        pass
+
+
+class RabbitPopulationDiffEq(TimeDependentDiffEq):
+    """
+    A simple differential equation modelling the growth of a rabbit population over time.
+    """
+
+    def __init__(self, n_0, r):
+        self.n_0 = n_0
+        self.r = r
+
+    def has_exact_solution(self) -> bool:
+        return True
+
+    def exact_y(self, t: float) -> float:
+        return self.n_0 * math.exp(self.r * t)
+
+    def y_0(self) -> float:
+        return self.n_0
+
+    def d_y(self, t: float, y: float) -> float:
+        return self.r * y

### 2020-03-23 code/python/src/parareal.py
Create a class for the Parareal framework

@@ -1,83 +1,96 @@
 import numpy as np
-import math
 
 from mpi4py import MPI
 
-from runge_kutta import ForwardEulerMethod, RK4
-
-r = .01
-n_0 = 10000.
-
-fine_d_t = .25
-coarse_d_t = 2 * fine_d_t
-k = 2
-
-g = ForwardEulerMethod()
-f = RK4()
-
-comm = MPI.COMM_WORLD
-rank = comm.Get_rank()
-size = comm.Get_size()
-
-duration = 8.
-time_slices = np.linspace(0., duration, size + 1)
+from diff_eq import TimeDependentDiffEq
+from runge_kutta import RungeKuttaMethod
 
 
-def operator(integration_method, y_0, t_0, t_max, d_t, d_y_wrt_t):
-    y = [y_0]
-    y_i = y_0
-    for i, t in enumerate(np.arange(t_0, t_max, d_t)):
-        t_i = t_0 + i * d_t
-        y_i = integration_method(y_i, t_i, d_t, d_y_wrt_t)
-        y.append(y_i)
-    return y[-1]
+class Parareal:
+    """
+    A parallel-in-time differential equation solver framework based on the Parareal algorithm.
+    """
 
+    def __init__(
+            self,
+            diff_eq: TimeDependentDiffEq,
+            fine_integrator: RungeKuttaMethod,
+            coarse_integrator: RungeKuttaMethod,
+            fine_d_t: float,
+            coarse_d_t: float,
+            t_max: float,
+            k: int):
+        self.diff_eq = diff_eq
+        self.fine_integrator = fine_integrator
+        self.coarse_integrator = coarse_integrator
+        self.fine_d_t = fine_d_t
+        self.coarse_d_t = coarse_d_t
+        self.t_max = t_max
+        self.k = k
 
-def calculate_corrections(y_0, t_0, t_max, d_y_wrt_t):
-    f_value = operator(f, y_0, t_0, t_max, fine_d_t, d_y_wrt_t)
-    g_value = operator(g, y_0, t_0, t_max, coarse_d_t, d_y_wrt_t)
-    return f_value - g_value
+    def _integrate(self, integrator: RungeKuttaMethod, y_0: float, t_0: float, t_max: float, d_t: float):
+        y = [y_0]
+        y_i = y_0
+        for i, t in enumerate(np.arange(t_0, t_max, d_t)):
+            t_i = t_0 + i * d_t
[…]

### 2020-03-23 code/python/src/main.py
Add a main script for running the Parareal solver

@@ -1,0 +1,17 @@
+from diff_eq import RabbitPopulationDiffEq
+from runge_kutta import ForwardEulerMethod, RK4
+from parareal import Parareal
+
+
+diff_eq = RabbitPopulationDiffEq(10000., .01)
+
+fine_integrator = RK4()
+coarse_integrator = ForwardEulerMethod()
+
+fine_d_t = .25
+coarse_d_t = 2 * fine_d_t
+t_max = 8.
+k = 2
+
+parareal = Parareal(diff_eq, fine_integrator, coarse_integrator, fine_d_t, coarse_d_t, t_max, k)
+parareal.run()

### 2020-03-23 code/python/tests/test_runge_kutta.py
Use underscore for unused parameter

@@ -40,7 +40,7 @@ def test_rk4():
     x1 = 1.
     d_x = 1.
 
-    def d_y_wrt_x(x, y): return 2 * y + 1
+    def d_y_wrt_x(_, y): return 2 * y + 1
 
     expected_y2 = 3.
     actual_y2 = rk4.integrate(y1, x1, d_x, d_y_wrt_x)

### 2020-03-23 code/python/src/main.py
Fix imports and the MPI launch command

@@ -1,6 +1,6 @@
-from diff_eq import RabbitPopulationDiffEq
-from runge_kutta import ForwardEulerMethod, RK4
-from parareal import Parareal
+from src.diff_eq import RabbitPopulationDiffEq
+from src.runge_kutta import ForwardEulerMethod, RK4
+from src.parareal import Parareal
 
 
 diff_eq = RabbitPopulationDiffEq(10000., .01)

### 2020-03-23 code/python/src/parareal.py

@@ -2,8 +2,8 @@ import numpy as np
 
 from mpi4py import MPI
 
-from diff_eq import TimeDependentDiffEq
-from runge_kutta import RungeKuttaMethod
+from src.diff_eq import TimeDependentDiffEq
+from src.runge_kutta import RungeKuttaMethod
 
 
 class Parareal:

### 2020-03-23 code/python/tests/test_runge_kutta.py

@@ -1,6 +1,6 @@
 import pytest
 
-from runge_kutta import ForwardEulerMethod, ExplicitMidpointMethod, RK4
+from src.runge_kutta import ForwardEulerMethod, ExplicitMidpointMethod, RK4
 
 
 def test_forward_euler_method():

### 2020-03-23 code/python/src/main.py
Add timing

@@ -3,14 +3,13 @@ from src.runge_kutta import ForwardEulerMethod, RK4
 from src.parareal import Parareal
 
 
-diff_eq = RabbitPopulationDiffEq(10000., .01)
-
+diff_eq = RabbitPopulationDiffEq(10000., 5e-5)
 fine_integrator = RK4()
 coarse_integrator = ForwardEulerMethod()
 
-fine_d_t = .25
+fine_d_t = .025
 coarse_d_t = 2 * fine_d_t
-t_max = 8.
+t_max = 100000.
 k = 2
 
 parareal = Parareal(diff_eq, fine_integrator, coarse_integrator, fine_d_t, coarse_d_t, t_max, k)

### 2020-03-23 code/python/src/parareal.py

@@ -74,6 +74,9 @@ class Parareal:
 
         y_coarse = np.copy(y)
 
+        comm.Barrier()
+        start_time = MPI.Wtime()
+
         for i in range(self.k):
             my_y_corr = self._calculate_corrections(y[rank], time_slices[rank], time_slices[rank + 1])
             comm.Allgather([my_y_corr, MPI.DOUBLE], [y_corr, MPI.DOUBLE])
@@ -81,6 +84,9 @@ class Parareal:
             for j, t in enumerate(time_slices[:-1]):
                 g_value = self.g(y[j], t, time_slices[j + 1])[-1]
                 y[j + 1] = g_value + y_corr[j]
+
+        comm.Barrier()
+        end_time = MPI.Wtime()
 
         if rank == 0:
             print('Coarse solution\n', y_coarse)
@@ -94,3 +100,5 @@ class Parareal:
                     y_exact[i + 1] = self.diff_eq.exact_y(t)
 
                 print('Analytic solution\n', y_exact)
+
+            print(f'Execution took {end_time - start_time}s')

### 2020-03-23 code/python/src/parareal.py
Retain the fine trajectory as the solution

@@ -29,18 +29,40 @@ class Parareal:
         self.k = k
 
     def _integrate(self, integrator: RungeKuttaMethod, y_0: float, t_0: float, t_max: float, d_t: float):
-        y = [y_0]
+        time_steps = np.arange(t_0, t_max, d_t)
+        y = np.empty(len(time_steps))
         y_i = y_0
         for i, t in enumerate(np.arange(t_0, t_max, d_t)):
             t_i = t_0 + i * d_t
             y_i = integrator.integrate(y_i, t_i, d_t, self.diff_eq.d_y)
-            y.append(y_i)
+            y[i] = y_i
         return y
 
-    def _calculate_corrections(self, y_0: float, t_0: float, t_max: float):
-        f_value = self.f(y_0, t_0, t_max)[-1]
-        g_value = self.g(y_0, t_0, t_max)[-1]
-        return f_value - g_value
+    def _print_results(self, time_slices, y_coarse, y, y_trajectory):
+        print('Coarse solution\n', y_coarse)
+        print('Fine solution\n', y)
+
+        print_trajectory = self.t_max / self.fine_d_t <= 50
+
+        if self.diff_eq.has_exact_solution():
+            y_exact = np.empty(len(time_slices))
+            y_exact[0] = self.diff_eq.y_0()
+
+            for i, t in enumerate(time_slices[1:]):
+                y_exact[i + 1] = self.diff_eq.exact_y(t)
+
+            print('Analytic solution\n', y_exact)
+
+            if print_trajectory:
+                y_exact_trajectory = np.empty(int(self.t_max / self.fine_d_t))
+
+                for i, t in enumerate(np.linspace(self.fine_d_t, self.t_max, len(y_exact_trajectory))):
+                    y_exact_trajectory[i] = self.diff_eq.exact_y(t)
+
+                print('Analytic trajectory:\n', y_exact_trajectory)
+
+        if print_trajectory:
+            print('Fine trajectory:\n', y_trajectory)
 
     """
     Returns the values of y between t_0 and t_max as estimated by the fine operator F.
@@ -62,43 +84,39 @@ class Parareal:
         rank = comm.Get_rank()
         size = comm.Get_size()
 
+        comm.barrier()
+        start_time = MPI.Wtime()
+
         time_slices = np.linspace(0., self.t_max, size + 1)
 
         y = np.empty(len(time_slices))
-        y_corr = np.empty(len(time_slices) - 1)
[…]

### 2020-03-23 code/python/src/parareal.py
Make sure at most n (number of processes) iterations are performed

@@ -99,7 +99,7 @@ class Parareal:
 
         y_coarse = np.copy(y) if rank == 0 else None
 
-        for i in range(self.k):
+        for i in range(min(size, self.k)):
             my_f_trajectory = self.f(y[rank], time_slices[rank], time_slices[rank + 1])
             my_g_trajectory = self.g(y[rank], time_slices[rank], time_slices[rank + 1])
 

### 2020-03-23 code/python/src/main.py
Update parameters

@@ -3,13 +3,13 @@ from src.runge_kutta import ForwardEulerMethod, RK4
 from src.parareal import Parareal
 
 
-diff_eq = RabbitPopulationDiffEq(10000., 5e-5)
+diff_eq = RabbitPopulationDiffEq(10000., 5e-2)
 fine_integrator = RK4()
 coarse_integrator = ForwardEulerMethod()
 
 fine_d_t = .025
 coarse_d_t = 2 * fine_d_t
-t_max = 100000.
+t_max = 1.
 k = 2
 
 parareal = Parareal(diff_eq, fine_integrator, coarse_integrator, fine_d_t, coarse_d_t, t_max, k)

### 2020-03-23 code/python/src/diff_eq.py
Rejig to ordinary differential equation and add x_0 and x_max methods

@@ -1,9 +1,10 @@
 import math
+from typing import Optional
 
 
-class TimeDependentDiffEq:
+class OrdinaryDiffEq:
     """
-    A representation of a time-dependent ordinary differential equation of the form y'(t) = f(t, y(t)).
+    A representation of an ordinary differential equation of the form y'(x) = f(x, y(x)).
     """
 
     """
@@ -13,41 +14,61 @@ class TimeDependentDiffEq:
         pass
 
     """
-    Returns the exact value of y(t) given t.
+    Returns the exact value of y(x) given x.
     """
-    def exact_y(self, t: float) -> float:
+    def exact_y(self, t: float) -> Optional[float]:
         pass
 
     """
-    Returns the value of y(0).
+    Returns the lower bound of the differential equation's solution's domain.
+    """
+    def x_0(self) -> float:
+        pass
+
+    """
+    Returns the upper bound of the differential equation's solution's domain.
+    """
+    def x_max(self) -> float:
+        pass
+
+    """
+    Returns the value of y(x_0).
     """
     def y_0(self) -> float:
         pass
 
     """
-    Returns the value of the derivative of y at t given t and y(t).
+    Returns the value of y'(x) given x and y(x).
     """
     def d_y(self, t: float, y: float) -> float:
         pass
 
 
-class RabbitPopulationDiffEq(TimeDependentDiffEq):
+class RabbitPopulationDiffEq(OrdinaryDiffEq):
     """
     A simple differential equation modelling the growth of a rabbit population over time.
     """
 
-    def __init__(self, n_0, r):
-        self.n_0 = n_0
-        self.r = r
+    def __init__(self, n_0, r, t_0, t_max):
+        self._n_0 = n_0
+        self._r = r
+        self._t_0 = t_0
+        self._t_max = t_max
 
     def has_exact_solution(self) -> bool:
         return True
 
     def exact_y(self, t: float) -> float:
-        return self.n_0 * math.exp(self.r * t)
+        return self._n_0 * math.exp(self._r * t)
+
+    def x_0(self) -> float:
+        return self._t_0
+
+    def x_max(self) -> float:
+        return self._t_max
 
     def y_0(self) -> float:
-        return self.n_0
+        return self._n_0
 
     def d_y(self, t: float, y: float) -> float:
-        return self.r * y
+        return self._r * y

### 2020-03-23 code/python/src/operator.py
Create a class for differential equation integration operators

@@ -1,0 +1,50 @@
+from typing import Sequence, Callable
+
+import numpy as np
+
+from src.runge_kutta import RungeKuttaMethod
+
+
+class Operator:
+    """
+    A base class for an operator to estimate the solution of a differential equation over a specific domain interval
+    given an initial value.
+    """
+
+    """
+    Returns an estimate of y(x_max).
+    """
+    def integrate(self, y_0: float, x_0: float, x_max: float, d_y: Callable[[float, float], float]) -> float:
+        pass
+
+
+class ConventionalOperator(Operator):
+    """
+    An operator using conventional Runge-Kutta integration.
+    """
+
+    def __init__(self, integrator: RungeKuttaMethod, d_x: float):
+        self._integrator = integrator
+        self._d_x = d_x
+
+    """
+    Returns the step size of the operator.
+    """
+    def d_x(self) -> float:
+        return self._d_x
+
+    """
+    Returns a discretised approximation of y over (x_0, x_max].
+    """
+    def trace(self, y_0: float, x_0: float, x_max: float, d_y: Callable[[float, float], float]) -> Sequence[float]:
+        x = np.arange(x_0, x_max, self._d_x)
+        y = np.empty(len(x))
+        y_i = y_0
+        for i, t in enumerate(x):
+            x_i = x_0 + i * self._d_x
+            y_i = self._integrator.integrate(y_i, x_i, self._d_x, d_y)
+            y[i] = y_i
+        return y
+
+    def integrate(self, y_0: float, x_0: float, x_max: float, d_y: Callable[[float, float], float]) -> float:
+        return self.trace(y_0, x_0, x_max, d_y)[-1]

### 2020-03-23 code/python/src/parareal.py
Parameterise the solver by the operators and return the solution

@@ -1,9 +1,11 @@
+from typing import Sequence
+
 import numpy as np
 
 from mpi4py import MPI
 
-from src.diff_eq import TimeDependentDiffEq
-from src.runge_kutta import RungeKuttaMethod
+from src.diff_eq import OrdinaryDiffEq
+from src.operator import Operator, ConventionalOperator
 
 
 class Parareal:
@@ -13,110 +15,84 @@ class Parareal:
 
     def __init__(
             self,
-            diff_eq: TimeDependentDiffEq,
-            fine_integrator: RungeKuttaMethod,
-            coarse_integrator: RungeKuttaMethod,
-            fine_d_t: float,
-            coarse_d_t: float,
-            t_max: float,
+            f: ConventionalOperator,
+            g: Operator,
             k: int):
-        self.diff_eq = diff_eq
-        self.fine_integrator = fine_integrator
-        self.coarse_integrator = coarse_integrator
-        self.fine_d_t = fine_d_t
-        self.coarse_d_t = coarse_d_t
-        self.t_max = t_max
+        self.f = f
+        self.g = g
         self.k = k
 
-    def _integrate(self, integrator: RungeKuttaMethod, y_0: float, t_0: float, t_max: float, d_t: float):
-        time_steps = np.arange(t_0, t_max, d_t)
-        y = np.empty(len(time_steps))
-        y_i = y_0
-        for i, t in enumerate(np.arange(t_0, t_max, d_t)):
-            t_i = t_0 + i * d_t
-            y_i = integrator.integrate(y_i, t_i, d_t, self.diff_eq.d_y)
-            y[i] = y_i
-        return y
-
-    def _print_results(self, time_slices, y_coarse, y, y_trajectory):
+    def _print_results(
+            self,
+            diff_eq: OrdinaryDiffEq,
+            time_slices: Sequence[float],
+            y_coarse: Sequence[float],
+            y: Sequence[float],
+            y_trajectory: Sequence[float]):
         print('Coarse solution\n', y_coarse)
         print('Fine solution\n', y)
 
-        print_trajectory = self.t_max / self.fine_d_t <= 50
+        print_trajectory = len(y_trajectory) <= 50
 
-        if self.diff_eq.has_exact_solution():
+        if diff_eq.has_exact_solution():
             y_exact = np.empty(len(time_slices))
-            y_exact[0] = self.diff_eq.y_0()
+            y_exact[0] = diff_eq.y_0()
 
[…]

### 2020-03-23 code/python/src/main.py
Update the construction and invocation of the solver

@@ -1,16 +1,11 @@
 from src.diff_eq import RabbitPopulationDiffEq
+from src.operator import ConventionalOperator
 from src.runge_kutta import ForwardEulerMethod, RK4
 from src.parareal import Parareal
 
 
-diff_eq = RabbitPopulationDiffEq(10000., 5e-2)
-fine_integrator = RK4()
-coarse_integrator = ForwardEulerMethod()
-
-fine_d_t = .025
-coarse_d_t = 2 * fine_d_t
-t_max = 1.
-k = 2
-
-parareal = Parareal(diff_eq, fine_integrator, coarse_integrator, fine_d_t, coarse_d_t, t_max, k)
-parareal.run()
+diff_eq = RabbitPopulationDiffEq(10000., 5e-2, 0., 1.)
+f = ConventionalOperator(RK4(), .025)
+g = ConventionalOperator(ForwardEulerMethod(), .1)
+solver = Parareal(f, g, 2)
+solver.solve(diff_eq)

### 2020-03-23 code/python/src/diff_eq.py
Update type hint

@@ -61,7 +61,7 @@ class RabbitPopulationDiffEq(OrdinaryDiffEq):
     def exact_y(self, t: float) -> float:
         return self._n_0 * math.exp(self._r * t)
 
-    def x_0(self) -> float:
+    def x_0(self) -> Optional[float]:
         return self._t_0
 
     def x_max(self) -> float:

### 2020-03-23 code/python/src/runge_kutta.py
Update documentation

@@ -7,7 +7,7 @@ class RungeKuttaMethod:
     """
 
     """
-    Estimates the value of y at x^ = x + d_x.
+    Estimates the value of y(x + d_x).
     """
     def integrate(self, y: float, x: float, d_x: float, d_y_wrt_x: Callable[[float, float], float]) -> float:
         pass

### 2020-03-23 code/python/src/integrator.py
Make distinction between operators and integrators clearer

@@ -1,9 +1,9 @@
 from typing import Callable
 
 
-class RungeKuttaMethod:
+class Integrator:
     """
-    A base class for Runge-Kutta methods
+    A base class for ordinary differential equation integrators.
     """
 
     """
@@ -13,7 +13,7 @@ class RungeKuttaMethod:
         pass
 
 
-class ForwardEulerMethod(RungeKuttaMethod):
+class ForwardEulerMethod(Integrator):
     """
     The forward Euler method, an explicit first order Runge-Kutta method.
     """
@@ -22,7 +22,7 @@ class ForwardEulerMethod(RungeKuttaMethod):
         return y + d_x * d_y_wrt_x(x, y)
 
 
-class ExplicitMidpointMethod(RungeKuttaMethod):
+class ExplicitMidpointMethod(Integrator):
     """
     The explicit midpoint method, a second order Runge-Kutta method.
     """
@@ -31,7 +31,7 @@ class ExplicitMidpointMethod(RungeKuttaMethod):
         return y + d_x * d_y_wrt_x(x + d_x / 2., y + d_y_wrt_x(x, y) * d_x / 2.)
 
 
-class RK4(RungeKuttaMethod):
+class RK4(Integrator):
     """
     The RK4 method, an explicit fourth order Runge-Kutta method.
     """

### 2020-03-23 code/python/src/main.py

@@ -1,6 +1,6 @@
 from src.diff_eq import RabbitPopulationDiffEq
 from src.operator import ConventionalOperator
-from src.runge_kutta import ForwardEulerMethod, RK4
+from src.integrator import ForwardEulerMethod, RK4
 from src.parareal import Parareal
 
 

### 2020-03-23 code/python/src/operator.py

@@ -1,8 +1,9 @@
-from typing import Sequence, Callable
+from typing import Sequence
 
 import numpy as np
 
-from src.runge_kutta import RungeKuttaMethod
+from src.diff_eq import OrdinaryDiffEq
+from src.integrator import Integrator
 
 
 class Operator:
@@ -12,18 +13,18 @@ class Operator:
     """
 
     """
-    Returns an estimate of y(x_max).
+    Returns an estimate of y(x_b) given x_a and y(x_a) for an x_b > x_a.
     """
-    def integrate(self, y_0: float, x_0: float, x_max: float, d_y: Callable[[float, float], float]) -> float:
+    def integrate(self, diff_eq: OrdinaryDiffEq, y_a: float, x_a: float, x_b: float) -> float:
         pass
 
 
 class ConventionalOperator(Operator):
     """
-    An operator using conventional Runge-Kutta integration.
+    An operator that uses conventional differential equation integration.
     """
 
-    def __init__(self, integrator: RungeKuttaMethod, d_x: float):
+    def __init__(self, integrator: Integrator, d_x: float):
         self._integrator = integrator
         self._d_x = d_x
 
@@ -34,17 +35,17 @@ class ConventionalOperator(Operator):
         return self._d_x
 
     """
-    Returns a discretised approximation of y over (x_0, x_max].
+    Returns a discretised approximation of y over (x_a, x_b].
     """
-    def trace(self, y_0: float, x_0: float, x_max: float, d_y: Callable[[float, float], float]) -> Sequence[float]:
-        x = np.arange(x_0, x_max, self._d_x)
+    def trace(self, diff_eq: OrdinaryDiffEq, y_a: float, x_a: float, x_b: float) -> Sequence[float]:
+        x = np.arange(x_a, x_b, self._d_x)
         y = np.empty(len(x))
-        y_i = y_0
+        y_i = y_a
         for i, t in enumerate(x):
-            x_i = x_0 + i * self._d_x
-            y_i = self._integrator.integrate(y_i, x_i, self._d_x, d_y)
+            x_i = x_a + i * self._d_x
+            y_i = self._integrator.integrate(y_i, x_i, self._d_x, diff_eq.d_y)
             y[i] = y_i
         return y
 
-    def integrate(self, y_0: float, x_0: float, x_max: float, d_y: Callable[[float, float], float]) -> float:
[…]

### 2020-03-23 code/python/src/parareal.py

@@ -70,19 +70,19 @@ class Parareal:
         y[0] = diff_eq.y_0()
 
         for i, t in enumerate(time_slices[:-1]):
-            y[i + 1] = self.g.integrate(y[i], t, time_slices[i + 1], diff_eq.d_y)
+            y[i + 1] = self.g.integrate(diff_eq, y[i], t, time_slices[i + 1])
         y_coarse = np.copy(y) if rank == 0 else None
 
         for i in range(min(size, self.k)):
-            my_f_trajectory = self.f.trace(y[rank], time_slices[rank], time_slices[rank + 1], diff_eq.d_y)
+            my_f_trajectory = self.f.trace(diff_eq, y[rank], time_slices[rank], time_slices[rank + 1])
             for j in range(size):
                 y_trajectory[j] = comm.bcast(my_f_trajectory, root=j)
 
-            my_g_value = self.g.integrate(y[rank], time_slices[rank], time_slices[rank + 1], diff_eq.d_y)
+            my_g_value = self.g.integrate(diff_eq, y[rank], time_slices[rank], time_slices[rank + 1])
             g_values = comm.allgather(my_g_value)
 
             for j, t in enumerate(time_slices[:-1]):
-                updated_g_value = self.g.integrate(y[j], t, time_slices[j + 1], diff_eq.d_y)
+                updated_g_value = self.g.integrate(diff_eq, y[j], t, time_slices[j + 1])
                 y[j + 1] = updated_g_value + y_trajectory[j][-1] - g_values[j]
                 y_trajectory[j] += updated_g_value - g_values[j]
 

### 2020-03-23 code/python/tests/test_integrator.py

@@ -1,6 +1,6 @@
 import pytest
 
-from src.runge_kutta import ForwardEulerMethod, ExplicitMidpointMethod, RK4
+from src.integrator import ForwardEulerMethod, ExplicitMidpointMethod, RK4
 
 
 def test_forward_euler_method():

### 2020-03-23 code/python/src/diff_eq.py
Add value checks

@@ -50,6 +50,7 @@ class RabbitPopulationDiffEq(OrdinaryDiffEq):
     """
 
     def __init__(self, n_0, r, t_0, t_max):
+        assert t_max > t_0
         self._n_0 = n_0
         self._r = r
         self._t_0 = t_0

### 2020-03-23 code/python/src/operator.py

@@ -38,6 +38,7 @@ class ConventionalOperator(Operator):
     Returns a discretised approximation of y over (x_a, x_b].
     """
     def trace(self, diff_eq: OrdinaryDiffEq, y_a: float, x_a: float, x_b: float) -> Sequence[float]:
+        assert x_b > x_a
         x = np.arange(x_a, x_b, self._d_x)
         y = np.empty(len(x))
         y_i = y_a

### 2020-03-24 code/python/src/diff_eq.py
Fix parameter names

@@ -16,7 +16,7 @@ class OrdinaryDiffEq:
     """
     Returns the exact value of y(x) given x.
     """
-    def exact_y(self, t: float) -> Optional[float]:
+    def exact_y(self, x: float) -> Optional[float]:
         pass
 
     """
@@ -40,7 +40,7 @@ class OrdinaryDiffEq:
     """
     Returns the value of y'(x) given x and y(x).
     """
-    def d_y(self, t: float, y: float) -> float:
+    def d_y(self, x: float, y: float) -> float:
         pass
 
 
@@ -59,8 +59,8 @@ class RabbitPopulationDiffEq(OrdinaryDiffEq):
     def has_exact_solution(self) -> bool:
         return True
 
-    def exact_y(self, t: float) -> float:
-        return self._n_0 * math.exp(self._r * t)
+    def exact_y(self, x: float) -> float:
+        return self._n_0 * math.exp(self._r * x)
 
     def x_0(self) -> Optional[float]:
         return self._t_0

### 2020-03-24 code/python/src/operator.py
Remove the integrate method and add an ML operator

@@ -1,4 +1,4 @@
-from typing import Sequence
+from typing import Sequence, Optional
 
 import numpy as np
 
@@ -13,9 +13,16 @@ class Operator:
     """
 
     """
-    Returns an estimate of y(x_b) given x_a and y(x_a) for an x_b > x_a.
+    Returns the step size of the operator. It may return None if the concept of step size does not make sense in the 
+    context of the operator.
     """
-    def integrate(self, diff_eq: OrdinaryDiffEq, y_a: float, x_a: float, x_b: float) -> float:
+    def d_x(self) -> Optional[float]:
+        pass
+
+    """
+    Returns a discretised approximation of y over (x_a, x_b]. It may return a sequence containing only y(x_b).
+    """
+    def trace(self, diff_eq: OrdinaryDiffEq, y_a: float, x_a: float, x_b: float) -> Sequence[float]:
         pass
 
 
@@ -28,25 +35,58 @@ class ConventionalOperator(Operator):
         self._integrator = integrator
         self._d_x = d_x
 
-    """
-    Returns the step size of the operator.
-    """
-    def d_x(self) -> float:
+    def d_x(self) -> Optional[float]:
         return self._d_x
 
-    """
-    Returns a discretised approximation of y over (x_a, x_b].
-    """
     def trace(self, diff_eq: OrdinaryDiffEq, y_a: float, x_a: float, x_b: float) -> Sequence[float]:
         assert x_b > x_a
         x = np.arange(x_a, x_b, self._d_x)
         y = np.empty(len(x))
         y_i = y_a
-        for i, t in enumerate(x):
-            x_i = x_a + i * self._d_x
+        for i, x_i in enumerate(x):
             y_i = self._integrator.integrate(y_i, x_i, self._d_x, diff_eq.d_y)
             y[i] = y_i
         return y
 
-    def integrate(self, diff_eq: OrdinaryDiffEq, y_a: float, x_a: float, x_b: float) -> float:
-        return self.trace(diff_eq, y_a, x_a, x_b)[-1]
+
+class MLOperator(Operator):
+    """
+    A machine learning accelerated operator that uses a regression model to integrate differential equations.
+    """
+
+    def __init__(self, model, trainer: ConventionalOperator, d_x: float):
+        self._model = model
+        self._trainer = trainer
+        self._d_x = d_x
[…]

### 2020-03-24 code/python/src/parareal.py

@@ -5,7 +5,7 @@ import numpy as np
 from mpi4py import MPI
 
 from src.diff_eq import OrdinaryDiffEq
-from src.operator import Operator, ConventionalOperator
+from src.operator import Operator
 
 
 class Parareal:
@@ -15,7 +15,7 @@ class Parareal:
 
     def __init__(
             self,
-            f: ConventionalOperator,
+            f: Operator,
             g: Operator,
             k: int):
         self.f = f
@@ -70,7 +70,7 @@ class Parareal:
         y[0] = diff_eq.y_0()
 
         for i, t in enumerate(time_slices[:-1]):
-            y[i + 1] = self.g.integrate(diff_eq, y[i], t, time_slices[i + 1])
+            y[i + 1] = self.g.trace(diff_eq, y[i], t, time_slices[i + 1])[-1]
         y_coarse = np.copy(y) if rank == 0 else None
 
         for i in range(min(size, self.k)):
@@ -78,11 +78,11 @@ class Parareal:
             for j in range(size):
                 y_trajectory[j] = comm.bcast(my_f_trajectory, root=j)
 
-            my_g_value = self.g.integrate(diff_eq, y[rank], time_slices[rank], time_slices[rank + 1])
+            my_g_value = self.g.trace(diff_eq, y[rank], time_slices[rank], time_slices[rank + 1])[-1]
             g_values = comm.allgather(my_g_value)
 
             for j, t in enumerate(time_slices[:-1]):
-                updated_g_value = self.g.integrate(diff_eq, y[j], t, time_slices[j + 1])
+                updated_g_value = self.g.trace(diff_eq, y[j], t, time_slices[j + 1])[-1]
                 y[j + 1] = updated_g_value + y_trajectory[j][-1] - g_values[j]
                 y_trajectory[j] += updated_g_value - g_values[j]
 

### 2020-03-24 code/python/src/main.py
Use a linear regression based ML operator as G

@@ -1,11 +1,14 @@
+from sklearn.linear_model import LinearRegression
+
 from src.diff_eq import RabbitPopulationDiffEq
-from src.operator import ConventionalOperator
-from src.integrator import ForwardEulerMethod, RK4
+from src.operator import ConventionalOperator, MLOperator
+from src.integrator import ExplicitMidpointMethod, RK4
 from src.parareal import Parareal
 
 
-diff_eq = RabbitPopulationDiffEq(10000., 5e-2, 0., 1.)
+diff_eq = RabbitPopulationDiffEq(10000., 5e-2, 0., 100.)
 f = ConventionalOperator(RK4(), .025)
-g = ConventionalOperator(ForwardEulerMethod(), .1)
+g_trainer = ConventionalOperator(ExplicitMidpointMethod(), .1)
+g = MLOperator(LinearRegression(), g_trainer, .5)
 solver = Parareal(f, g, 2)
 solver.solve(diff_eq)

### 2020-03-24 code/python/src/operator.py
Format code

@@ -1,4 +1,4 @@
-from typing import Sequence, Optional
+from typing import Sequence
 
 import numpy as np
 
@@ -13,14 +13,13 @@ class Operator:
     """
 
     """
-    Returns the step size of the operator. It may return None if the concept of step size does not make sense in the 
-    context of the operator.
+    Returns the step size of the operator.
     """
-    def d_x(self) -> Optional[float]:
+    def d_x(self) -> float:
         pass
 
     """
-    Returns a discretised approximation of y over (x_a, x_b]. It may return a sequence containing only y(x_b).
+    Returns a discretised approximation of y over (x_a, x_b].
     """
     def trace(self, diff_eq: OrdinaryDiffEq, y_a: float, x_a: float, x_b: float) -> Sequence[float]:
         pass
@@ -35,17 +34,18 @@ class ConventionalOperator(Operator):
         self._integrator = integrator
         self._d_x = d_x
 
-    def d_x(self) -> Optional[float]:
+    def d_x(self) -> float:
         return self._d_x
 
     def trace(self, diff_eq: OrdinaryDiffEq, y_a: float, x_a: float, x_b: float) -> Sequence[float]:
-        assert x_b > x_a
         x = np.arange(x_a, x_b, self._d_x)
         y = np.empty(len(x))
         y_i = y_a
+
         for i, x_i in enumerate(x):
             y_i = self._integrator.integrate(y_i, x_i, self._d_x, diff_eq.d_y)
             y[i] = y_i
+
         return y
 
 
@@ -68,25 +68,29 @@ class MLOperator(Operator):
         obs = np.empty((len(x) - 1, 2))
         y = np.empty(len(obs))
         y_i = diff_eq.y_0()
+
         for i, x_i in enumerate(x[:-1]):
             obs[i][0] = y_i
             obs[i][1] = diff_eq.d_y(x_i, y_i)
             y_i = self._trainer.trace(diff_eq, y_i, x_i, x[i + 1])[-1]
             y[i] = y_i
+
         self._model.fit(obs, y)
         self._diff_eq_trained_on = diff_eq
 
-    def d_x(self) -> Optional[float]:
+    def d_x(self) -> float:
         return self._d_x
 
     def trace(self, diff_eq: OrdinaryDiffEq, y_a: float, x_a: float, x_b: float) -> Sequence[float]:
-        assert x_b > x_a
         if diff_eq != self._diff_eq_trained_on:
[…]

### 2020-03-24 report/MSc-report.tex

@@
 \renewcommand{\theenumii}{\theenumi.\arabic{enumii}.}
-% this will make nested numbering all arabic numbers
-
-% This example file shows how a report can be laid out using Latex. It
-% does not use any special local features so should be portable to other
-% places.
-% 
-% When producing draft copies of a thesis you may want to print only
-% selected pages of the thesis. To do this use the command
-% 
-% dvips -f -p 4 -n 3 myfile.dvi | lpr
-% 
-% where -p 4 means start printing at page 4 (ie the page that will be
-% numbered 4, not necessarily the 4th page) and -n 3 means print 3 pages.
-% This example will print pages 4, 5 and 6.
-% 
-% If you want to print the report and also save paper you can print more
-% than one page on each sheet of paper. Use the command
-% 
-% dvips -f myfile.dvi | psnup -2 | lpr
-% 
-% to print 2 pages per sheet. psnup can take values 2, 4, 8, or 9.
-%
-% To produce a PDF version you can create a PostScript copy first
-%
-% dvips -f myfile.dvi > myfile.pdf
-%
-% and then convert it
-%
-% distill myfile.ps
-%
-% or you can go straight to PDF
-%
-% pdflatex myfile
-%
-% Note that pdflatex expects all included figures to be in PDF too. See
-% the includegraphics command below.
-
-% This document contains many cross-references and forward references,
-% eg in constructing a table of contents, so Latex may need to be run
-% twice to get all the references correct. If you need to run Latex twice
-% you may get the warning:
-% 
-% LaTeX Warning: Label(s) may have changed. Rerun to get cross-references right.
 
@@
 
-\title{A \LaTeX{}  Report Example}
-\author{David Henty \& MSc Dissertations Team}
+\title{Parallel-in-time Methods with Machine Learning - Report}
+\author{Viktor Csomor}
 \date{\today}
@@
 \tableofcontents
-
-\newpage
-
-\section*{Acknowledgements}
-
-This template is a slightly modified version of the one developed by
-Prof. Charles Duncan for MSc students in the Dept. of Meteorology. His
-acknowledgement follows:
-
-{\em This template has been produced with help from many former students who
-have shown different ways of doing things. Please make suggestions for
-further improvements.}
-
-\section*{Note}
-
-This template was originally created for a full MSc dissertation.
-Although I have modified the format slightly (eg there are no separate
-chapters), I have not substantially changed the text or the advice
-appearing in the \LaTeX comments (source lines beginning with \%). You
-should therefore be aware that for a shorter report you may
-not need to go into as much detail as indicated here. For example, you
-might not need to include any acknowledgements and any reference list is
-likely to be very short.
 

### 2020-03-24 report/MSc-report.tex
Update on Overleaf.

@@
 is no substitute for thorough proof reading!}
+
+Some dynamical systems have an inherently finite spatial domain and there are only so many subdomains this space can be decomposed into.
+The number of these subdomains limits the level of parallelism achievable using traditional parallel numerical methods.
+Parallel-in-time methods address this by solving the differential equation describing the system in parallel along the time axis unlocking potentially higher levels of parallelism.
+The most notable parallel-in-time method, the Parareal algorithm, does this by using an expensive fine operator and a computationally cheap coarse operator
 

### 2020-03-24 code/python/src/diff_eq.py
Fix type hint

@@ -59,10 +59,10 @@ class RabbitPopulationDiffEq(OrdinaryDiffEq):
     def has_exact_solution(self) -> bool:
         return True
 
-    def exact_y(self, x: float) -> float:
+    def exact_y(self, x: float) -> Optional[float]:
         return self._n_0 * math.exp(self._r * x)
 
-    def x_0(self) -> Optional[float]:
+    def x_0(self) -> float:
         return self._t_0
 
     def x_max(self) -> float:

### 2020-03-24 code/python/src/main.py
Update model training process

@@ -9,6 +9,6 @@ from src.parareal import Parareal
 diff_eq = RabbitPopulationDiffEq(10000., 5e-2, 0., 100.)
 f = ConventionalOperator(RK4(), .025)
 g_trainer = ConventionalOperator(ExplicitMidpointMethod(), .1)
-g = MLOperator(LinearRegression(), g_trainer, .5)
+g = MLOperator(LinearRegression(), g_trainer, .5, 20)
 solver = Parareal(f, g, 2)
 solver.solve(diff_eq)

### 2020-03-24 code/python/src/operator.py

@@ -1,4 +1,4 @@
-from typing import Sequence
+from typing import Sequence, Any
 
 import numpy as np
 
@@ -54,10 +54,11 @@ class MLOperator(Operator):
     A machine learning accelerated operator that uses a regression model to integrate differential equations.
     """
 
-    def __init__(self, model, trainer: ConventionalOperator, d_x: float):
+    def __init__(self, model: Any, trainer: Operator, d_x: float, data_epochs: int):
         self._model = model
         self._trainer = trainer
         self._d_x = d_x
+        self._data_epochs = data_epochs
         self._diff_eq_trained_on = None
 
     """
@@ -65,15 +66,20 @@ class MLOperator(Operator):
     """
     def train_model(self, diff_eq: OrdinaryDiffEq):
         x = np.arange(diff_eq.x_0(), diff_eq.x_max() + self._d_x / 2, self._d_x)
-        obs = np.empty((len(x) - 1, 2))
+        obs = np.empty((self._data_epochs * (len(x) - 1), 3))
         y = np.empty(len(obs))
-        y_i = diff_eq.y_0()
 
-        for i, x_i in enumerate(x[:-1]):
-            obs[i][0] = y_i
-            obs[i][1] = diff_eq.d_y(x_i, y_i)
-            y_i = self._trainer.trace(diff_eq, y_i, x_i, x[i + 1])[-1]
-            y[i] = y_i
+        for k in range(self._data_epochs):
+            offset = k * (len(x) - 1)
+            y_i = diff_eq.y_0()
+            for i, x_i in enumerate(x[:-1]):
+                ind = offset + i
+                d_y_i = diff_eq.d_y(x_i, y_i)
+                obs[ind][0] = x_i
+                obs[ind][1] = y_i
+                obs[ind][2] = d_y_i
+                y[ind] = self._trainer.trace(diff_eq, y_i, x_i, x[i + 1])[-1]
+                y_i = y[ind] + np.random.normal(0., d_y_i * self._d_x)
 
         self._model.fit(obs, y)
         self._diff_eq_trained_on = diff_eq
@@ -90,7 +96,7 @@ class MLOperator(Operator):
         y_i = y_a
 
         for i, x_i in enumerate(x):
-            y_i = self._model.predict([[y_i, diff_eq.d_y(x_i, y_i)]])[0]
+            y_i = self._model.predict([[x_i, y_i, diff_eq.d_y(x_i, y_i)]])[0]
             y[i] = y_i
 
         return y

### 2020-03-24 code/python/src/main.py
Time the serial solver and the parallel solvers with and without ML

@@ -1,14 +1,28 @@
+from mpi4py import MPI
 from sklearn.linear_model import LinearRegression
 
 from src.diff_eq import RabbitPopulationDiffEq
+from src.integrator import ForwardEulerMethod, RK4
 from src.operator import ConventionalOperator, MLOperator
-from src.integrator import ExplicitMidpointMethod, RK4
 from src.parareal import Parareal
 
+diff_eq = RabbitPopulationDiffEq(1000., 1e-4, 0., 50000.)
+f = ConventionalOperator(RK4(), .01)
+g = ConventionalOperator(ForwardEulerMethod(), .1)
+k = 1
 
-diff_eq = RabbitPopulationDiffEq(10000., 5e-2, 0., 100.)
-f = ConventionalOperator(RK4(), .025)
-g_trainer = ConventionalOperator(ExplicitMidpointMethod(), .1)
-g = MLOperator(LinearRegression(), g_trainer, .5, 20)
-solver = Parareal(f, g, 2)
+solver = Parareal(f, g, k)
 solver.solve(diff_eq)
+
+ml_g = MLOperator(LinearRegression(), g, 100., 10)
+ml_g.train_model(diff_eq)
+
+ml_solver = Parareal(f, ml_g, k)
+ml_solver.solve(diff_eq)
+
+if MPI.COMM_WORLD.rank == 0:
+    start_time = MPI.Wtime()
+    y_x_max = f.trace(diff_eq, diff_eq.y_0(), diff_eq.x_0(), diff_eq.x_max())[-1]
+    end_time = MPI.Wtime()
+    print(f'Serial fine solution: {y_x_max}')
+    print(f'Serial execution took {end_time - start_time}s')

### 2020-03-24 code/python/src/diff_eq.py
Fix formatting and documentation

@@ -4,49 +4,53 @@ from typing import Optional
 
 class OrdinaryDiffEq:
     """
-    A representation of an ordinary differential equation of the form y'(x) = f(x, y(x)).
+    A representation of an ordinary differential equation of the form
+    y'(x) = f(x, y(x)).
     """
 
-    """
-    Returns whether the differential equation has an analytic solution
-    """
     def has_exact_solution(self) -> bool:
+        """
+        Returns whether the differential equation has an analytic solution
+        """
         pass
 
-    """
-    Returns the exact value of y(x) given x.
-    """
     def exact_y(self, x: float) -> Optional[float]:
+        """
+        Returns the exact value of y(x) given x.
+        """
         pass
 
-    """
-    Returns the lower bound of the differential equation's solution's domain.
-    """
     def x_0(self) -> float:
+        """
+        Returns the lower bound of the differential equation's solution's
+        domain.
+        """
         pass
 
-    """
-    Returns the upper bound of the differential equation's solution's domain.
-    """
     def x_max(self) -> float:
+        """
+        Returns the upper bound of the differential equation's solution's
+        domain.
+        """
         pass
 
-    """
-    Returns the value of y(x_0).
-    """
     def y_0(self) -> float:
+        """
+        Returns the value of y(x_0).
+        """
         pass
 
-    """
-    Returns the value of y'(x) given x and y(x).
-    """
     def d_y(self, x: float, y: float) -> float:
+        """
+        Returns the value of y'(x) given x and y(x).
+        """
         pass
 
 
 class RabbitPopulationDiffEq(OrdinaryDiffEq):
     """
-    A simple differential equation modelling the growth of a rabbit population over time.
+    A simple differential equation modelling the growth of a rabbit population
+    over time.
     """
 
     def __init__(self, n_0, r, t_0, t_max):

### 2020-03-24 code/python/src/integrator.py

@@ -6,10 +6,21 @@ class Integrator:
     A base class for ordinary differential equation integrators.
     """
 
-    """
-    Estimates the value of y(x + d_x).
-    """
-    def integrate(self, y: float, x: float, d_x: float, d_y_wrt_x: Callable[[float, float], float]) -> float:
+    def integrate(
+            self,
+            y: float,
+            x: float,
+            d_x: float,
+            d_y_wrt_x: Callable[[float, float], float]) -> float:
+        """
+        Estimates the value of y(x + d_x).
+
+        :param y: the value of y(x)
+        :param x: the value of x
+        :param d_x: the amount of increase in x
+        :param d_y_wrt_x: the value of y'(x)
+        :return: the value of y(x + d_x).
+        """
         pass
 
 
@@ -18,7 +29,12 @@ class ForwardEulerMethod(Integrator):
     The forward Euler method, an explicit first order Runge-Kutta method.
     """
 
-    def integrate(self, y: float, x: float, d_x: float, d_y_wrt_x: Callable[[float, float], float]) -> float:
+    def integrate(
+            self,
+            y: float,
+            x: float,
+            d_x: float,
+            d_y_wrt_x: Callable[[float, float], float]) -> float:
         return y + d_x * d_y_wrt_x(x, y)
 
 
@@ -27,8 +43,15 @@ class ExplicitMidpointMethod(Integrator):
     The explicit midpoint method, a second order Runge-Kutta method.
     """
 
-    def integrate(self, y: float, x: float, d_x: float, d_y_wrt_x: Callable[[float, float], float]) -> float:
-        return y + d_x * d_y_wrt_x(x + d_x / 2., y + d_y_wrt_x(x, y) * d_x / 2.)
+    def integrate(
+            self,
+            y: float,
+            x: float,
+            d_x: float,
+            d_y_wrt_x: Callable[[float, float], float]) -> float:
+        return y + d_x * d_y_wrt_x(
+            x + d_x / 2.,
+            y + d_y_wrt_x(x, y) * d_x / 2.)
 
 
 class RK4(Integrator):
@@ -36,7 +59,12 @@ class RK4(Integrator):
     The RK4 method, an explicit fourth order Runge-Kutta method.
     """
 
-    def integrate(self, y: float, x: float, d_x: float, d_y_wrt_x: Callable[[float, float], float]) -> float:
+    def integrate(
+            self,
[…]

### 2020-03-24 code/python/src/main.py

@@ -22,7 +22,9 @@ ml_solver.solve(diff_eq)
 
 if MPI.COMM_WORLD.rank == 0:
     start_time = MPI.Wtime()
-    y_x_max = f.trace(diff_eq, diff_eq.y_0(), diff_eq.x_0(), diff_eq.x_max())[-1]
+    y_x_max = f.trace(
+        diff_eq, diff_eq.y_0(), diff_eq.x_0(), diff_eq.x_max())[-1]
     end_time = MPI.Wtime()
+
     print(f'Serial fine solution: {y_x_max}')
     print(f'Serial execution took {end_time - start_time}s')

### 2020-03-24 code/python/src/operator.py

@@ -8,20 +8,36 @@ from src.integrator import Integrator
 
 class Operator:
     """
-    A base class for an operator to estimate the solution of a differential equation over a specific domain interval
-    given an initial value.
+    A base class for an operator to estimate the solution of a differential
+    equation over a specific domain interval given an initial value.
     """
 
-    """
-    Returns the step size of the operator.
-    """
     def d_x(self) -> float:
+        """
+        Returns the step size of the operator.
+        """
         pass
 
-    """
-    Returns a discretised approximation of y over (x_a, x_b].
-    """
-    def trace(self, diff_eq: OrdinaryDiffEq, y_a: float, x_a: float, x_b: float) -> Sequence[float]:
+    def trace(
+            self,
+            diff_eq: OrdinaryDiffEq,
+            y_a: float,
+            x_a: float,
+            x_b: float) -> Sequence[float]:
+        """
+        Returns a discretised approximation of y over (x_a, x_b].
+
+        :param diff_eq: the differential equation whose solution's trajectory
+        is to be traced
+        :param y_a: y(x_a), that is the value of the differential equation's
+        solution at the lower bound of the interval it is to be traced over
+        :param x_a: the lower bound of the interval over which the differential
+        equation's solution is to be traced (exclusive)
+        :param x_b: the upper bound of the interval over which the differential
+        equation's solution is to be traced (inclusive)
+        :return: a sequence of floating points number representing the
+        discretised solution of the differential equation y over (x_a, x_b]
+        """
         pass
 
 
@@ -37,7 +53,12 @@ class ConventionalOperator(Operator):
     def d_x(self) -> float:
         return self._d_x
 
-    def trace(self, diff_eq: OrdinaryDiffEq, y_a: float, x_a: float, x_b: float) -> Sequence[float]:
+    def trace(
+            self,
+            diff_eq: OrdinaryDiffEq,
+            y_a: float,
+            x_a: float,
+            x_b: float) -> Sequence[float]:
         x = np.arange(x_a, x_b, self._d_x)
         y = np.empty(len(x))
         y_i = y_a
@@ -51,21 +72,25 @@ class ConventionalOperator(Operator):
 
 class MLOperator(Operator):
     """
-    A machine learning accelerated operator that uses a regression model to integrate differential equations.
+    A machine learning accelerated operator that uses a regression model to
[…]

### 2020-03-24 code/python/src/parareal.py

@@ -10,7 +10,8 @@ from src.operator import Operator
 
 class Parareal:
     """
-    A parallel-in-time differential equation solver framework based on the Parareal algorithm.
+    A parallel-in-time differential equation solver framework based on the
+    Parareal algorithm.
     """
 
     def __init__(
@@ -46,17 +47,25 @@ class Parareal:
                 y_exact_trajectory = np.empty(len(y_trajectory))
 
                 for i, t in enumerate(
-                        np.linspace(diff_eq.x_0() + self.f.d_x(), diff_eq.x_max(), len(y_exact_trajectory))):
+                        np.linspace(
+                            diff_eq.x_0() + self.f.d_x(),
+                            diff_eq.x_max(),
+                            len(y_exact_trajectory))):
                     y_exact_trajectory[i] = diff_eq.exact_y(t)
                 print('Analytic trajectory:\n', y_exact_trajectory)
 
         if print_trajectory:
             print('Fine trajectory:\n', y_trajectory)
 
-    """
-    Runs the Parareal solver and returns the discretised solution of the differential equation.
-    """
     def solve(self, diff_eq: OrdinaryDiffEq) -> Sequence[float]:
+        """
+        Runs the Parareal solver and returns the discretised solution of the
+        differential equation.
+
+        :param diff_eq: the differential equation to solve
+        :return: the discretised trajectory of the differential equation's
+        solution
+        """
         comm = MPI.COMM_WORLD
         comm.barrier()
         start_time = MPI.Wtime()
@@ -66,7 +75,8 @@ class Parareal:
         time_slices = np.linspace(diff_eq.x_0(), diff_eq.x_max(), size + 1)
 
         y = np.empty(len(time_slices))
-        y_trajectory = np.empty((size, int(time_slices[-1] / (size * self.f.d_x()))))
+        y_trajectory = np.empty(
+            (size, int(time_slices[-1] / (size * self.f.d_x()))))
         y[0] = diff_eq.y_0()
 
         for i, t in enumerate(time_slices[:-1]):
@@ -74,15 +84,18 @@ class Parareal:
         y_coarse = np.copy(y) if rank == 0 else None
 
         for i in range(min(size, self.k)):
-            my_f_trajectory = self.f.trace(diff_eq, y[rank], time_slices[rank], time_slices[rank + 1])
+            my_f_trajectory = self.f.trace(
[…]

### 2020-03-25 code/python/src/diff_eq.py
Add more documentation

@@ -54,6 +54,12 @@ class RabbitPopulationDiffEq(OrdinaryDiffEq):
     """
 
     def __init__(self, n_0, r, t_0, t_max):
+        """
+        :param n_0: the initial population size
+        :param r: the population growth rate
+        :param t_0: the start time
+        :param t_max: the end time
+        """
         assert t_max > t_0
         self._n_0 = n_0
         self._r = r

### 2020-03-25 code/python/src/operator.py

@@ -47,6 +47,10 @@ class ConventionalOperator(Operator):
     """
 
     def __init__(self, integrator: Integrator, d_x: float):
+        """
+        :param integrator: the differential equation integrator to use
+        :param d_x: the step size to use with the integrator
+        """
         self._integrator = integrator
         self._d_x = d_x
 
@@ -78,6 +82,20 @@ class MLOperator(Operator):
 
     def __init__(
             self, model: Any, trainer: Operator, d_x: float, data_epochs: int):
+        """
+        :param model: the regression model to use as the integrator; its input
+        are the values of x, y(x), and y'(x) and its output is y(x + d_x) where
+        d_x is the step size of this operator defined by the corresponding
+        constructor argument
+        :param trainer: the operator for generating the labels for the data
+        to train the regression model on
+        :param d_x: the step size of the operator; it determines the lengths of
+        the domain slices over which the training operator is used to trace the
+        differential equation's solution and provide the labels for the
+        training data
+        :param data_epochs: the number of iterations to perform over the domain
+        of the differential equation to generate the training data
+        """
         self._model = model
         self._trainer = trainer
         self._d_x = d_x
@@ -88,6 +106,19 @@ class MLOperator(Operator):
         """
         Trains the regression model behind the operator on the provided
         differential equation.
+
+        It generates the training data by repeatedly iterating over the domain
+        of the differential equation in steps of size d_x and tracing the
+        solution using the training operator. At every step i, a new training
+        data point is created out of the values of x_i, y(x_i), and y'(x_i)
+        labelled by y(x_i+1) = y(x_i + d_x) as estimated by the training
+        operator. Once the data point is created, a 0-mean Gaussian noise is
+        added to the value of y(x_i+1) to perturbate the trajectory of y.
+        This introduces some variance to the training data and helps better
+        approximate the function represented by the training operator. The
+        standard deviation of this Gaussian is y'(x_i) * d_x.
+
+        :param diff_eq: the differential equation to train the model on
         """
         x = np.arange(
             diff_eq.x_0(), diff_eq.x_max() + self._d_x / 2, self._d_x)

### 2020-03-25 code/python/src/parareal.py

@@ -19,6 +19,12 @@ class Parareal:
             f: Operator,
             g: Operator,
             k: int):
+        """
+        :param f: the fine operator
+        :param g: the coarse operator
+        :param k: the number of corrective iterations to perform using the fine
+        operator. It is capped at the number of processes running the solver.
+        """
         self.f = f
         self.g = g
         self.k = k
@@ -57,12 +63,17 @@ class Parareal:
         if print_trajectory:
             print('Fine trajectory:\n', y_trajectory)
 
-    def solve(self, diff_eq: OrdinaryDiffEq) -> Sequence[float]:
+    def solve(
+            self,
+            diff_eq: OrdinaryDiffEq,
+            verbose: bool = True) -> Sequence[float]:
         """
         Runs the Parareal solver and returns the discretised solution of the
         differential equation.
 
         :param diff_eq: the differential equation to solve
+        :param verbose: whether information about the solution and the
+        execution time should be printed
         :return: the discretised trajectory of the differential equation's
         solution
         """
@@ -104,7 +115,7 @@ class Parareal:
         comm.barrier()
         end_time = MPI.Wtime()
 
-        if rank == 0:
+        if verbose and rank == 0:
             self._print_results(
                 diff_eq, time_slices, y_coarse, y, y_trajectory)
             print(f'Execution took {end_time - start_time}s')

### 2020-03-25 code/python/src/main.py
Move timing and result analysis out of the parareal class

@@ -6,25 +6,46 @@ from src.integrator import ForwardEulerMethod, RK4
 from src.operator import ConventionalOperator, MLOperator
 from src.parareal import Parareal
 
-diff_eq = RabbitPopulationDiffEq(1000., 1e-4, 0., 50000.)
-f = ConventionalOperator(RK4(), .01)
-g = ConventionalOperator(ForwardEulerMethod(), .1)
-k = 1
 
-solver = Parareal(f, g, k)
-solver.solve(diff_eq)
-
-ml_g = MLOperator(LinearRegression(), g, 100., 10)
-ml_g.train_model(diff_eq)
-
-ml_solver = Parareal(f, ml_g, k)
-ml_solver.solve(diff_eq)
-
-if MPI.COMM_WORLD.rank == 0:
+def time_parallel_solver_and_print_result(parallel_solver, solver_name):
+    comm.barrier()
     start_time = MPI.Wtime()
-    y_x_max = f.trace(
+    y_max = parallel_solver.solve(diff_eq)[-1]
+    comm.barrier()
+    end_time = MPI.Wtime()
+    if comm.rank == 0:
+        print(f'{solver_name} solution: {y_max}; '
+              f'execution took {end_time - start_time}s')
+
+
+def time_operator_and_print_result(operator, operator_name):
+    start_time = MPI.Wtime()
+    y_max = operator.trace(
         diff_eq, diff_eq.y_0(), diff_eq.x_0(), diff_eq.x_max())[-1]
     end_time = MPI.Wtime()
+    print(f'{operator_name} solution: {y_max}; '
+          f'execution took {end_time - start_time}s')
 
-    print(f'Serial fine solution: {y_x_max}')
-    print(f'Serial execution took {end_time - start_time}s')
+
+comm = MPI.COMM_WORLD
+
+diff_eq = RabbitPopulationDiffEq(1000., 1e-4, 0., 50000.)
+
+f = ConventionalOperator(RK4(), .01)
+g = ConventionalOperator(ForwardEulerMethod(), .1)
+g_ml = MLOperator(LinearRegression(), g, 100., 10)
+
+k = 1
+
+parareal_ml = Parareal(f, g_ml, k)
+parareal = Parareal(f, g, k)
+
+time_parallel_solver_and_print_result(parareal_ml, 'Parareal ML w/ training')
+time_parallel_solver_and_print_result(parareal_ml, 'Parareal ML w/o training')
[…]

### 2020-03-25 code/python/src/parareal.py

@@ -29,79 +29,43 @@ class Parareal:
         self.g = g
         self.k = k
 
-    def _print_results(
-            self,
-            diff_eq: OrdinaryDiffEq,
-            time_slices: Sequence[float],
-            y_coarse: Sequence[float],
-            y: Sequence[float],
-            y_trajectory: Sequence[float]):
-        print('Coarse solution\n', y_coarse)
-        print('Fine solution\n', y)
-
-        print_trajectory = len(y_trajectory) <= 50
-
-        if diff_eq.has_exact_solution():
-            y_exact = np.empty(len(time_slices))
-            y_exact[0] = diff_eq.y_0()
-
-            for i, t in enumerate(time_slices[1:]):
-                y_exact[i + 1] = diff_eq.exact_y(t)
-            print('Analytic solution\n', y_exact)
-
-            if print_trajectory:
-                y_exact_trajectory = np.empty(len(y_trajectory))
-
-                for i, t in enumerate(
-                        np.linspace(
-                            diff_eq.x_0() + self.f.d_x(),
-                            diff_eq.x_max(),
-                            len(y_exact_trajectory))):
-                    y_exact_trajectory[i] = diff_eq.exact_y(t)
-                print('Analytic trajectory:\n', y_exact_trajectory)
-
-        if print_trajectory:
-            print('Fine trajectory:\n', y_trajectory)
-
-    def solve(
-            self,
-            diff_eq: OrdinaryDiffEq,
-            verbose: bool = True) -> Sequence[float]:
+    def solve(self, diff_eq: OrdinaryDiffEq) -> Sequence[float]:
         """
         Runs the Parareal solver and returns the discretised solution of the
         differential equation.
 
         :param diff_eq: the differential equation to solve
-        :param verbose: whether information about the solution and the
-        execution time should be printed
         :return: the discretised trajectory of the differential equation's
         solution
         """
         comm = MPI.COMM_WORLD
-        comm.barrier()
-        start_time = MPI.Wtime()
 
-        rank = comm.Get_rank()
-        size = comm.Get_size()
-        time_slices = np.linspace(diff_eq.x_0(), diff_eq.x_max(), size + 1)
-
+        time_slices = np.linspace(
+            diff_eq.x_0(), diff_eq.x_max(), comm.size + 1)
         y = np.empty(len(time_slices))
         y_trajectory = np.empty(
[…]

### 2020-03-25 code/python/src/main.py
Improve the readability of the parareal implementation

@@ -45,7 +45,7 @@ time_parallel_solver_and_print_result(parareal_ml, 'Parareal ML w/o training')
 time_parallel_solver_and_print_result(parareal, 'Parareal')
 
 if comm.rank == 0:
-    print(f'Analytic solution: {diff_eq.exact_y(diff_eq.x_max())}')
-    time_operator_and_print_result(f, 'Fine')
     time_operator_and_print_result(g, 'Coarse')
     time_operator_and_print_result(g_ml, 'Coarse ML')
+    time_operator_and_print_result(f, 'Fine')
+    print(f'Analytic solution: {diff_eq.exact_y(diff_eq.x_max())}')

### 2020-03-25 code/python/src/parareal.py

@@ -69,9 +69,14 @@ class Parareal:
             g_values = comm.allgather(my_g_value)
 
             for j, t in enumerate(time_slices[:-1]):
-                updated_g_value = self.g.trace(
-                    diff_eq, y[j], t, time_slices[j + 1])[-1]
-                y[j + 1] = updated_g_value + y_trajectory[j][-1] - g_values[j]
-                y_trajectory[j] += updated_g_value - g_values[j]
+                g_value = g_values[j]
+                new_g_value = self.g.trace(
+                    diff_eq,
+                    y[j],
+                    t,
+                    time_slices[j + 1])[-1]
+
+                y_trajectory[j] += new_g_value - g_value
+                y[j + 1] = y_trajectory[j][-1]
 
         return y_trajectory.flatten()

### 2020-03-25 code/python/src/operator.py
Address potential floating point issues

@@ -12,6 +12,15 @@ class Operator:
     equation over a specific domain interval given an initial value.
     """
 
+    def _discretise_sub_domain(self, x_a: float, x_b: float):
+        """
+        Returns a discretisation of the the interval [x_a, x_b^) using the
+        operator's step size, d_x, where x_b^ is x_b rounded to the nearest
+        multiple of d_x.
+        """
+        adjusted_x_b = self.d_x() * round(x_b / self.d_x())
+        return np.arange(x_a, adjusted_x_b, self.d_x())
+
     def d_x(self) -> float:
         """
         Returns the step size of the operator.
@@ -63,7 +72,7 @@ class ConventionalOperator(Operator):
             y_a: float,
             x_a: float,
             x_b: float) -> Sequence[float]:
-        x = np.arange(x_a, x_b, self._d_x)
+        x = self._discretise_sub_domain(x_a, x_b)
         y = np.empty(len(x))
         y_i = y_a
 
@@ -120,21 +129,21 @@ class MLOperator(Operator):
 
         :param diff_eq: the differential equation to train the model on
         """
-        x = np.arange(
-            diff_eq.x_0(), diff_eq.x_max() + self._d_x / 2, self._d_x)
-        obs = np.empty((self._data_epochs * (len(x) - 1), 3))
+        x = self._discretise_sub_domain(diff_eq.x_0(), diff_eq.x_max())
+        obs = np.empty((self._data_epochs * len(x), 3))
         y = np.empty(len(obs))
 
         for k in range(self._data_epochs):
-            offset = k * (len(x) - 1)
+            offset = k * len(x)
             y_i = diff_eq.y_0()
-            for i, x_i in enumerate(x[:-1]):
+            for i, x_i in enumerate(x):
                 ind = offset + i
                 d_y_i = diff_eq.d_y(x_i, y_i)
                 obs[ind][0] = x_i
                 obs[ind][1] = y_i
                 obs[ind][2] = d_y_i
-                y[ind] = self._trainer.trace(diff_eq, y_i, x_i, x[i + 1])[-1]
+                y[ind] = self._trainer.trace(
+                    diff_eq, y_i, x_i, x_i + self._d_x)[-1]
                 y_i = y[ind] + np.random.normal(0., d_y_i * self._d_x)
 
         self._model.fit(obs, y)
@@ -152,7 +161,7 @@ class MLOperator(Operator):
[…]

### 2020-03-25 code/python/src/parareal.py
Use more efficient communication

@@ -42,31 +42,30 @@ class Parareal:
 
         time_slices = np.linspace(
             diff_eq.x_0(), diff_eq.x_max(), comm.size + 1)
-        y = np.empty(len(time_slices))
         y_trajectory = np.empty(
             (comm.size, int(time_slices[-1] / (comm.size * self.f.d_x()))))
+        g_values = np.empty(comm.size)
+        y = np.empty(len(time_slices))
         y[0] = diff_eq.y_0()
 
         for i, t in enumerate(time_slices[:-1]):
             y[i + 1] = self.g.trace(diff_eq, y[i], t, time_slices[i + 1])[-1]
 
         for i in range(min(comm.size, self.k)):
-            my_f_trajectory = self.f.trace(
+            my_y_trajectory = self.f.trace(
                 diff_eq,
                 y[comm.rank],
                 time_slices[comm.rank],
                 time_slices[comm.rank + 1])
-
-            for j in range(comm.size):
-                y_trajectory[j] = comm.bcast(my_f_trajectory, root=j)
+            comm.Allgather(
+                [my_y_trajectory, MPI.DOUBLE], [y_trajectory, MPI.DOUBLE])
 
             my_g_value = self.g.trace(
                 diff_eq,
                 y[comm.rank],
                 time_slices[comm.rank],
                 time_slices[comm.rank + 1])[-1]
-
-            g_values = comm.allgather(my_g_value)
+            comm.Allgather([my_g_value, MPI.DOUBLE], [g_values, MPI.DOUBLE])
 
             for j, t in enumerate(time_slices[:-1]):
                 g_value = g_values[j]

### 2020-03-25 code/python/src/parareal.py
Prepend an underscore to each member variable's name

@@ -25,9 +25,9 @@ class Parareal:
         :param k: the number of corrective iterations to perform using the fine
         operator. It is capped at the number of processes running the solver.
         """
-        self.f = f
-        self.g = g
-        self.k = k
+        self._f = f
+        self._g = g
+        self._k = k
 
     def solve(self, diff_eq: OrdinaryDiffEq) -> Sequence[float]:
         """
@@ -43,16 +43,16 @@ class Parareal:
         time_slices = np.linspace(
             diff_eq.x_0(), diff_eq.x_max(), comm.size + 1)
         y_trajectory = np.empty(
-            (comm.size, int(time_slices[-1] / (comm.size * self.f.d_x()))))
+            (comm.size, int(time_slices[-1] / (comm.size * self._f.d_x()))))
         g_values = np.empty(comm.size)
         y = np.empty(len(time_slices))
         y[0] = diff_eq.y_0()
 
         for i, t in enumerate(time_slices[:-1]):
-            y[i + 1] = self.g.trace(diff_eq, y[i], t, time_slices[i + 1])[-1]
+            y[i + 1] = self._g.trace(diff_eq, y[i], t, time_slices[i + 1])[-1]
 
-        for i in range(min(comm.size, self.k)):
-            my_y_trajectory = self.f.trace(
+        for i in range(min(comm.size, self._k)):
+            my_y_trajectory = self._f.trace(
                 diff_eq,
                 y[comm.rank],
                 time_slices[comm.rank],
@@ -60,7 +60,7 @@ class Parareal:
             comm.Allgather(
                 [my_y_trajectory, MPI.DOUBLE], [y_trajectory, MPI.DOUBLE])
 
-            my_g_value = self.g.trace(
+            my_g_value = self._g.trace(
                 diff_eq,
                 y[comm.rank],
                 time_slices[comm.rank],
@@ -69,7 +69,7 @@ class Parareal:
 
             for j, t in enumerate(time_slices[:-1]):
                 g_value = g_values[j]
-                new_g_value = self.g.trace(
+                new_g_value = self._g.trace(
                     diff_eq,
                     y[j],
                     t,

### 2020-03-25 code/python/src/parareal.py
Gather y trajectory only once at the end

@@ -42,14 +42,17 @@ class Parareal:
 
         time_slices = np.linspace(
             diff_eq.x_0(), diff_eq.x_max(), comm.size + 1)
-        y_trajectory = np.empty(
-            (comm.size, int(time_slices[-1] / (comm.size * self._f.d_x()))))
-        g_values = np.empty(comm.size)
         y = np.empty(len(time_slices))
         y[0] = diff_eq.y_0()
 
+        f_values = np.empty(comm.size)
+        g_values = np.empty(comm.size)
+        new_g_values = np.empty(comm.size)
+
         for i, t in enumerate(time_slices[:-1]):
             y[i + 1] = self._g.trace(diff_eq, y[i], t, time_slices[i + 1])[-1]
+
+        my_y_trajectory = None
 
         for i in range(min(comm.size, self._k)):
             my_y_trajectory = self._f.trace(
@@ -57,8 +60,9 @@ class Parareal:
                 y[comm.rank],
                 time_slices[comm.rank],
                 time_slices[comm.rank + 1])
+            my_f_value = my_y_trajectory[-1]
             comm.Allgather(
-                [my_y_trajectory, MPI.DOUBLE], [y_trajectory, MPI.DOUBLE])
+                [my_f_value, MPI.DOUBLE], [f_values, MPI.DOUBLE])
 
             my_g_value = self._g.trace(
                 diff_eq,
@@ -68,14 +72,23 @@ class Parareal:
             comm.Allgather([my_g_value, MPI.DOUBLE], [g_values, MPI.DOUBLE])
 
             for j, t in enumerate(time_slices[:-1]):
+                f_value = f_values[j]
                 g_value = g_values[j]
+                correction = f_value - g_value
+
                 new_g_value = self._g.trace(
                     diff_eq,
                     y[j],
                     t,
                     time_slices[j + 1])[-1]
+                new_g_values[j] = new_g_value
 
-                y_trajectory[j] += new_g_value - g_value
-                y[j + 1] = y_trajectory[j][-1]
+                y[j + 1] = new_g_value + correction
 
-        return y_trajectory.flatten()
+        my_y_trajectory += new_g_values[comm.rank] - g_values[comm.rank]
+        y_trajectory = np.empty(
+            comm.size * int(time_slices[-1] / (comm.size * self._f.d_x())))
+        comm.Allgather(
+            [my_y_trajectory, MPI.DOUBLE], [y_trajectory, MPI.DOUBLE])
+
+        return y_trajectory

### 2020-03-25 code/python/src/operator.py
Add missing type hints

@@ -1,4 +1,4 @@
-from typing import Sequence, Any
+from typing import Sequence, Any, Optional
 
 import numpy as np
 
@@ -12,7 +12,10 @@ class Operator:
     equation over a specific domain interval given an initial value.
     """
 
-    def _discretise_sub_domain(self, x_a: float, x_b: float):
+    def _discretise_sub_domain(
+            self,
+            x_a: float,
+            x_b: float) -> Sequence[float]:
         """
         Returns a discretisation of the the interval [x_a, x_b^) using the
         operator's step size, d_x, where x_b^ is x_b rounded to the nearest
@@ -109,7 +112,7 @@ class MLOperator(Operator):
         self._trainer = trainer
         self._d_x = d_x
         self._data_epochs = data_epochs
-        self._diff_eq_trained_on = None
+        self._diff_eq_trained_on: Optional[OrdinaryDiffEq] = None
 
     def train_model(self, diff_eq: OrdinaryDiffEq):
         """

### 2020-03-26 report/MSc-report.tex
Update on Overleaf.

--- 2020-03-24 report/MSc-report.tex
@@
 
-This should contain a description of your project and the problem you
-are trying to solve. Where appropriate you should also include
-references to work which has already been done on your topic and
-anything else which lets you set your work in context.
+Differential equations can model a wide array of dynamical systems that play important roles in science, engineering, and finance.
+Examples of these include population growth, the motion of fluids, and changes in stock prices.
+While some simple differential equations can be solved analytically, most of them require a numerical approach \cite[p.~310]{suli2003}.
+Due to the importance of these models, their accurate and efficient numerical solution is a well researched area of numerical analysis.
+A lot of this research is focused on the parallelisation of numerical solvers to achieve better accuracy or reduce execution times.
+Traditionally, this parallelisation is applied either across the system or across time \cite{gear1988}.
 
-One of the things you will need to do is to ensure that you have a
-suitable list of references.  To do this you should see \cite{ref:lam}
-or some other suitable reference.  Note the format of the citation used
-here is the style favoured in this department.  Here is another
-reference \cite{ref:bloggs} for good measure.
+\section{Background and Literature Review}
 
-You will also want to make sure you have no spelling or grammatical
-mistakes. To help idwentify spelling mistukes you caan use the commands
-{\em ispell} or {\em spell}. See the appropriate manual pages. Remember
-that spelling mistakes are not the only errors which can occur. Spelling
-checkers will not find errors which are, in fact, valid words such as
-{\em there} for {\em their}, nor will they find repeated words which
-sometimes occur if your concentration is broken when typing. {\bf There
-is no substitute for thorough proof reading!}
+To do...
+
+\section{Preliminary Investigations}
+
+\section{Final Proposal}
+
+\section{Workplan}
+
+\begin{enumerate}
+    \item Implement framework
+        \begin{enumerate}
+        	\item Add support for systems of ODEs
+        	\item Add support for PDEs
+        	\item Add all the ML operators
+        	\item Optimise
+        \end{enumerate}
+    \item Benchmark
+        \begin{enumerate}
+        	\item Compare the ML operator to conventional ones
+        	\item Compare the Parareal framework with and without ML
+        	\item Measure the execution time of the Parareal (ML) framework against a traditional parallel solver
+        \end{enumerate}
+    \item Write dissertation
+\end{enumerate}
+
+\section{Risk Analysis}
+
+\begin{itemize}
+    \item Failure to implement a single framework to handle both ODEs and PDEs
+    \item Failure to fit a regression model to complex differential equations
+    \item Failure to utilise transfer learning
+    \item Unavailability of computing resources
+    \item Disruptions due to Covid-19
+\end{itemize}
+
+\section{Outline of the Dissertation Report}
+
+\begin{enumerate}
+	\item Introduction
+	\item Background
+    	\begin{enumerate}
+    		\item Parareal algorithm
+    		\item Machine learning for differential equations
+    		\item Parareal with neural network based fine operator
+    	\end{enumerate}
+	\item Problem formulation
+    	\begin{enumerate}
+    		\item Operator-agnostic Parareal framework
+    		\item Advantages of machine learning for coarse operator
+    		\item Transfer learning
+    	\end{enumerate}
+	\item Design and implementation
+    	\begin{enumerate}
+    	    \item Language and tools
+    		\item Parareal framework
+            	\begin{enumerate}
+            		\item Differential equations
+            		\item Integrators
+            		\item Operators
+            	\end{enumerate}
+    		\item Machine learning accelerated operators
+    	\end{enumerate}
+    \item Results
+        \begin{enumerate}
+    		\item Machine learning operator performance
+    		\item Parareal performance with and without machine learning
+    		\item Parareal performance against conventional parallel solver
+    	\end{enumerate}
+	\item Conclusions
+\end{enumerate}
 
@@
 
-In the introductory paragraph, clearly identify the reviewed dissertation by its Title, Author and Year.
-
-\subsection*{Summarize the project in your own words.}
-
-This is just to show how to break things into sections.
-
-Here is a padding paragraph.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
-Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
-more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
-More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More
-rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
-Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
-more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
-More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.   Rhubarb.  More
-rhubarb.  Yet more rhubarb.   Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.    Rhubarb.  More
-rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.
-
-\subsection*{Does the dissertation adequately explain the context in which the problem is set?}
-
-Give evidence to support your answer based on the quality of motivation, background and previous work overview.
-
-Here is a padding paragraph.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
-Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
-more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
-More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More
-rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
-Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
-more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
-More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.   Rhubarb.  More
-rhubarb.  Yet more rhubarb.   Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.    Rhubarb.  More
-rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.
-
-\subsection*{Has the student adequately explained the scope of the problem, the approach that they would take, how success would be measured and reflected on the outcomes?}
-
-Consider limitations, reproducibility and research methods used. Give evidence to support your answer.
-
-\subsection*{Quality of presentation.}
-Comment on the overall structure, whether the explanations were clear and easy to follow, how was the language used.
-
-\subsection*{Presentation of additional information.}
-Discus the use of references and citations throughout the text and use of figures and tables (their design, placement and descriptions, references to from within text, etc.)
-
-\section{Background and Literature Review}
-
-I created a figure which is centred and stretched to 30\% of the
-width of the page \verb+{0.30\hsize}+ and with the height stretched by
-the same amount \verb+{!}+ to preserve the aspect ratio. If you omit the
-extension (ie .eps, .ps or .pdf) on the file name then \LaTeX  will pick
-up the postscript copy whereas pdflatex will automatically pick up the
-PDF version.
-
-\begin{figure}
-
-\begin{center}
-\resizebox{0.30\hsize}{!}{\includegraphics{logos/crest_bw}}
-\end{center}
-
-\caption{The University Crest}
-\label{fig:eucrest}
-
-\end{figure}
-
-% see the man page for dvips for details of the special command which is
-% much more powerful than is shown here. It allows offsets in the
-% horizontal and vertical and scaling in x and y.
-
-% choosing suitable values for offset and scale can be a tiring matter
-% of trial and error.
-
-% note that labels do not need to include a description of the object
-% they are labelling but it can be helpful, eg \label{fig:figurename}.
-
-You can use a label on a figure to refer to it later: The university
-crest is in Figure~\ref{fig:eucrest}. Note that you should not use phrases like
-``the figure above'' or ``the following figure'' since \LaTeX  may move
-the figure relative to the text if it cannot be fitted onto the current page.
-
-\section{Preliminary Investigations}
-
-You might want to include an equation here:
-
-\begin{equation} \delta N_{\nu} = (\delta N_{\nu})_{ex} + (\delta N_{\nu})_{au} 
-\label{equation:delsplit}
-% note that the label is optional but allows you to refer to this
-% equation later.
-\end{equation}
-
-You might sometimes want to include equations without numbering them.
-
-\begin{displaymath}
-E=mc^{2}
-\end{displaymath}
-
-You might also want to include diagrams.  The example shows the use of
-the special command which allows existing postscript files to be
-included.  You would normally keep your figures separate from the text. 
-These pictures might be satellite images or postscript output from some
-program such as IDL, PV-WAVE, Uniras or xpaint.
-
-\paragraph{More on numbering:}
-
-This text is in a paragraph which is also not numbered by default and
-the ``title'' of the paragraph is not on a separate line.
-If you want to increase the depth to which subsections are numbered you
-should see the subsection on setting the secnumdepth counter in the manual.
-
-\section{Final Proposal}
-
-One way to produce simple graphs is to use gnuplot which can produce
-native \LaTeX  output. Graph \ref{fig:gnu} was produced using gnuplot with
-output designated as \LaTeX; this creates a \LaTeX output file which you
-can include directly or keep separate and refer to using the {\em
-include} command.
-
-Another approach is to draw simple figures using {\em xfig} which allows
-you to export diagrams in \LaTeX picture format so that the diagram can
-be included directly.
-
-Perhaps the most robust way to include graphs is to create then using
-whatever package you like (eg Excel) and then convert them to PostScript
-or PDF and include them in the same was as was done in
-Figure~\ref{fig:eucrest} for the University Crest. You can usually do
-this with most packages, including Microsoft ones. One trick for
-producing PostScript is to print to a dummy PostScript printer; for PDF
-you can use the free CutePDF utilities at {\verb+www.cutepdf.com+}.
-
-% in practice you would probably keep this in a separate file and use
-% the \include{filename} command to insert it here.
-
-\begin{figure}
-% GNUPLOT: LaTeX picture
-\setlength{\unitlength}{0.240900pt}
-\ifx\plotpoint\undefined\newsavebox{\plotpoint}\fi
-\sbox{\plotpoint}{\rule[-0.200pt]{0.400pt}{0.400pt}}%
-\begin{picture}(1500,1200)(0,0)
-\font\gnuplot=cmr10 at 10pt
-\gnuplot
-\sbox{\plotpoint}{\rule[-0.200pt]{0.400pt}{0.400pt}}%
-\put(220.0,113.0){\rule[-0.200pt]{292.934pt}{0.400pt}}
-\put(220.0,113.0){\rule[-0.200pt]{0.400pt}{245.477pt}}
-\put(220.0,113.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
-\put(198,113){\makebox(0,0)[r]{$0$}}
-\put(1416.0,113.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
-\put(220.0,317.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
-\put(198,317){\makebox(0,0)[r]{$0.2$}}
-\put(1416.0,317.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
-\put(220.0,521.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
-\put(198,521){\makebox(0,0)[r]{$0.4$}}
-\put(1416.0,521.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
-\put(220.0,724.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
-\put(198,724){\makebox(0,0)[r]{$0.6$}}
-\put(1416.0,724.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
-\put(220.0,928.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
-\put(198,928){\makebox(0,0)[r]{$0.8$}}
-\put(1416.0,928.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
-\put(220.0,1132.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
-\put(198,1132){\makebox(0,0)[r]{$1$}}
-\put(1416.0,1132.0){\rule[-0.200pt]{4.818pt}{0.400pt}}
-\put(220.0,113.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
-\put(220,68){\makebox(0,0){$0$}}
-\put(220.0,1112.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
-\put(414.0,113.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
-\put(414,68){\makebox(0,0){$1$}}
-\put(414.0,1112.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
-\put(607.0,113.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
-\put(607,68){\makebox(0,0){$2$}}
-\put(607.0,1112.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
-\put(801.0,113.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
-\put(801,68){\makebox(0,0){$3$}}
-\put(801.0,1112.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
-\put(995.0,113.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
-\put(995,68){\makebox(0,0){$4$}}
-\put(995.0,1112.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
-\put(1188.0,113.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
-\put(1188,68){\makebox(0,0){$5$}}
-\put(1188.0,1112.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
-\put(1382.0,113.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
-\put(1382,68){\makebox(0,0){$6$}}
-\put(1382.0,1112.0){\rule[-0.200pt]{0.400pt}{4.818pt}}
-\put(220.0,113.0){\rule[-0.200pt]{292.934pt}{0.400pt}}
-\put(1436.0,113.0){\rule[-0.200pt]{0.400pt}{245.477pt}}
-\put(220.0,1132.0){\rule[-0.200pt]{292.934pt}{0.400pt}}
-\put(45,622){\makebox(0,0){\shortstack{This is\\the\\$y$ axis}}}
-\put(828,23){\makebox(0,0){This is the $x$ axis}}
-\put(828,1177){\makebox(0,0){This is a plot of $y=sin(x)$}}
-\put(220.0,113.0){\rule[-0.200pt]{0.400pt}{245.477pt}}
-\sbox{\plotpoint}{\rule[-0.500pt]{1.000pt}{1.000pt}}%
-\put(1306,1067){\makebox(0,0)[r]{sin(x)}}
-\multiput(1328,1067)(20.756,0.000){4}{\usebox{\plotpoint}}
-\put(1394,1067){\usebox{\plotpoint}}
-\put(220,113){\usebox{\plotpoint}}
-\multiput(220,113)(3.768,20.411){4}{\usebox{\plotpoint}}
-\multiput(232,178)(4.132,20.340){3}{\usebox{\plotpoint}}
-\multiput(245,242)(3.825,20.400){3}{\usebox{\plotpoint}}
-\multiput(257,306)(3.884,20.389){3}{\usebox{\plotpoint}}
-\multiput(269,369)(3.944,20.377){3}{\usebox{\plotpoint}}
-\multiput(281,431)(4.326,20.300){3}{\usebox{\plotpoint}}
-\multiput(294,492)(4.137,20.339){3}{\usebox{\plotpoint}}
-\multiput(306,551)(4.276,20.310){3}{\usebox{\plotpoint}}
-\multiput(318,608)(4.693,20.218){3}{\usebox{\plotpoint}}
-\multiput(331,664)(4.583,20.243){2}{\usebox{\plotpoint}}
-\multiput(343,717)(4.754,20.204){3}{\usebox{\plotpoint}}
-\multiput(355,768)(5.034,20.136){2}{\usebox{\plotpoint}}
-\multiput(367,816)(5.760,19.940){2}{\usebox{\plotpoint}}
-\multiput(380,861)(5.579,19.992){3}{\usebox{\plotpoint}}
-\put(398.00,923.50){\usebox{\plotpoint}}
-\multiput(404,943)(7.049,19.522){2}{\usebox{\plotpoint}}
-\multiput(417,979)(7.288,19.434){2}{\usebox{\plotpoint}}
-\put(433.18,1021.10){\usebox{\plotpoint}}
-\multiput(441,1040)(8.982,18.712){2}{\usebox{\plotpoint}}
-\put(460.41,1076.97){\usebox{\plotpoint}}
-\put(471.84,1094.28){\usebox{\plotpoint}}
-\put(484.84,1110.41){\usebox{\plotpoint}}
-\put(500.42,1124.01){\usebox{\plotpoint}}
-\multiput(503,1126)(19.159,7.983){0}{\usebox{\plotpoint}}
-\put(519.48,1131.37){\usebox{\plotpoint}}
-\multiput(527,1132)(20.136,-5.034){0}{\usebox{\plotpoint}}
-\put(539.74,1128.60){\usebox{\plotpoint}}
-\put(557.04,1117.38){\usebox{\plotpoint}}
-\put(570.79,1101.95){\usebox{\plotpoint}}
-\put(582.44,1084.80){\usebox{\plotpoint}}
-\put(593.09,1066.99){\usebox{\plotpoint}}
-\multiput(601,1053)(8.430,-18.967){2}{\usebox{\plotpoint}}
-\put(619.18,1010.54){\usebox{\plotpoint}}
-\multiput(625,996)(7.413,-19.387){2}{\usebox{\plotpoint}}
-\multiput(638,962)(6.403,-19.743){2}{\usebox{\plotpoint}}
-\multiput(650,925)(5.830,-19.920){2}{\usebox{\plotpoint}}
-\multiput(662,884)(5.461,-20.024){2}{\usebox{\plotpoint}}
-\multiput(674,840)(5.533,-20.004){2}{\usebox{\plotpoint}}
-\multiput(687,793)(4.937,-20.160){3}{\usebox{\plotpoint}}
-\multiput(699,744)(4.667,-20.224){2}{\usebox{\plotpoint}}
-\multiput(711,692)(4.858,-20.179){3}{\usebox{\plotpoint}}
-\multiput(724,638)(4.276,-20.310){3}{\usebox{\plotpoint}}
-\multiput(736,581)(4.205,-20.325){3}{\usebox{\plotpoint}}
-\multiput(748,523)(4.070,-20.352){3}{\usebox{\plotpoint}}
-\multiput(760,463)(4.326,-20.300){3}{\usebox{\plotpoint}}
-\multiput(773,402)(3.884,-20.389){3}{\usebox{\plotpoint}}
-\multiput(785,339)(3.884,-20.389){3}{\usebox{\plotpoint}}
-\multiput(797,276)(4.070,-20.352){3}{\usebox{\plotpoint}}
-\multiput(810,211)(3.825,-20.400){3}{\usebox{\plotpoint}}
-\multiput(822,147)(3.607,-20.440){2}{\usebox{\plotpoint}}
-\put(828,113){\usebox{\plotpoint}}
-\end{picture}
-\caption{Simple Gnuplot example}
-\label{fig:gnu}
-\end{figure}
-
-\section{Workplan}
-
-You are likely to use tables and graphs. You can create tables easily in \LaTeX.
-
-\begin{table}[h]
-\begin{center}
-\begin{tabular}{ l|c|l }
-\toprule
-{\bf File names} & {\bf Satellite} & {\bf Resolution}\\
-\midrule
-  worldr         &  Meteosat     &   5km\\
-  worldg         &  Meteosat     &   5km\\
-  worldb         &  Meteosat     &   5km\\
-\bottomrule
-\end{tabular}
-\end{center}
-\caption{This is a simple table. More complicated tables can have
-headings which pass over more than one column}
-\label{simple_table}
-\end{table}
-
-If you want to produce fancier tables than shown in Table \ref{simple_table}
-refer to the manual.
-
-\section{Risk Analysis}
-
-Here is a padding paragraph.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
-Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
-more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
-More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More
-rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.
-
-\section{Outline of the Dissertation Report}
-
-You will probably need a numbered list here, which is rather simple in \LaTeX:
-
-\begin{enumerate}
-	\item This is top level rhubarb.
-	\begin{enumerate}
-		\item 2nd level
-		\item some rhubarb
-		\item and more
-	\end{enumerate}
-	\item More top level rhubarb
-	\begin{enumerate}
-		\item other boring vegetables 
-		\item kale, for example
-	\end{enumerate}
-	\item And so on
-\end{enumerate}
+Fitting Large-Scale Gaussian Mixtures With Accelerated Gradient Descent, Nestor Sanchez, 2018.
 
 \pagebreak
-\begin{thebibliography}{100}
 
-\bibitem{ref:lam} L.Lamport. {\em 1986 Latex User's Guide
-and Reference Manual.} Addison Wesley. pp242.
-
-\bibitem{ref:bloggs} F.Bloggs. {\em 1993 Latex Users do it
-in Environments} Int. Journal of Silly Findings. pp 23-29.
-
-\end{thebibliography}
+\bibliographystyle{unsrt}
+\bibliography{ref}
 

### 2020-03-26 code/python/src/main.py
Use update magnitude to determine number of corrective iterations

@@ -2,7 +2,7 @@ from mpi4py import MPI
 from sklearn.linear_model import LinearRegression
 
 from src.diff_eq import RabbitPopulationDiffEq
-from src.integrator import ForwardEulerMethod, RK4
+from src.integrator import ExplicitMidpointMethod, RK4
 from src.operator import ConventionalOperator, MLOperator
 from src.parareal import Parareal
 
@@ -10,7 +10,7 @@ from src.parareal import Parareal
 def time_parallel_solver_and_print_result(parallel_solver, solver_name):
     comm.barrier()
     start_time = MPI.Wtime()
-    y_max = parallel_solver.solve(diff_eq)[-1]
+    y_max = parallel_solver.solve(diff_eq, threshold)[-1]
     comm.barrier()
     end_time = MPI.Wtime()
     if comm.rank == 0:
@@ -32,13 +32,13 @@ comm = MPI.COMM_WORLD
 diff_eq = RabbitPopulationDiffEq(1000., 1e-4, 0., 50000.)
 
 f = ConventionalOperator(RK4(), .01)
-g = ConventionalOperator(ForwardEulerMethod(), .1)
+g = ConventionalOperator(ExplicitMidpointMethod(), .25)
 g_ml = MLOperator(LinearRegression(), g, 100., 10)
 
-k = 1
+parareal = Parareal(f, g)
+parareal_ml = Parareal(f, g_ml)
 
-parareal_ml = Parareal(f, g_ml, k)
-parareal = Parareal(f, g, k)
+threshold = 1e-3
 
 time_parallel_solver_and_print_result(parareal_ml, 'Parareal ML w/ training')
 time_parallel_solver_and_print_result(parareal_ml, 'Parareal ML w/o training')

### 2020-03-26 code/python/src/parareal.py

@@ -17,24 +17,27 @@ class Parareal:
     def __init__(
             self,
             f: Operator,
-            g: Operator,
-            k: int):
+            g: Operator):
         """
         :param f: the fine operator
         :param g: the coarse operator
-        :param k: the number of corrective iterations to perform using the fine
-        operator. It is capped at the number of processes running the solver.
         """
         self._f = f
         self._g = g
-        self._k = k
 
-    def solve(self, diff_eq: OrdinaryDiffEq) -> Sequence[float]:
+    def solve(
+            self,
+            diff_eq: OrdinaryDiffEq,
+            threshold: float) -> Sequence[float]:
         """
         Runs the Parareal solver and returns the discretised solution of the
         differential equation.
 
         :param diff_eq: the differential equation to solve
+        :param threshold: the minimum absolute value of the largest update to
+        the solution required to perform another corrective iteration; if all
+        updates are smaller than the threshold, the solution is considered
+        accurate enough
         :return: the discretised trajectory of the differential equation's
         solution
         """
@@ -54,7 +57,7 @@ class Parareal:
 
         my_y_trajectory = None
 
-        for i in range(min(comm.size, self._k)):
+        for i in range(comm.size):
             my_y_trajectory = self._f.trace(
                 diff_eq,
                 y[comm.rank],
@@ -71,6 +74,8 @@ class Parareal:
                 time_slices[comm.rank + 1])[-1]
             comm.Allgather([my_g_value, MPI.DOUBLE], [g_values, MPI.DOUBLE])
 
+            max_update = 0.
+
             for j, t in enumerate(time_slices[:-1]):
                 f_value = f_values[j]
                 g_value = g_values[j]
@@ -83,7 +88,13 @@ class Parareal:
                     time_slices[j + 1])[-1]
                 new_g_values[j] = new_g_value
 
-                y[j + 1] = new_g_value + correction
+                new_y_next = new_g_value + correction
+                max_update = max(max_update, abs(new_y_next - y[j + 1]))
+
+                y[j + 1] = new_y_next
+
+            if max_update < threshold:
+                break
 
         my_y_trajectory += new_g_values[comm.rank] - g_values[comm.rank]
         y_trajectory = np.empty(

### 2020-03-26 report/MSc-report.tex
Update on Overleaf.

@@
 \usepackage{booktabs}
+\usepackage{pgfgantt}
 
@@
 Differential equations can model a wide array of dynamical systems that play important roles in science, engineering, and finance.
-Examples of these include population growth, the motion of fluids, and changes in stock prices.
+Examples of these include population growth, the motion of fluids, and stock market dynamics.
 While some simple differential equations can be solved analytically, most of them require a numerical approach \cite[p.~310]{suli2003}.
 Due to the importance of these models, their accurate and efficient numerical solution is a well researched area of numerical analysis.
-A lot of this research is focused on the parallelisation of numerical solvers to achieve better accuracy or reduce execution times.
-Traditionally, this parallelisation is applied either across the system or across time \cite{gear1988}.
+A lot of this research is focused on the parallelisation of numerical solvers to achieve higher accuracy or reduce execution times.
+Usually, this parallelisation is applied either across the system or across time \cite{gear1988}.
+The former generally means exploiting parallelism for the evaluation or integration of the right-hand side of a differential equation at a single point \cite[p.~1-2]{solodushkin2016}.
+While it is a straightforward approach, it works well only for high dimensional differential equations.
+On the other hand, parallel-in-time methods allow for the utilisation of high levels of parallelism regardless the dimensionality of the problem.
+
+Our project focuses on the Parareal algorithm \cite{parareal} from the family of parallel-in-time methods.
+This algorithm divides the time domain up into $P$ slices where $P$ is the number of executing processes or threads.
 
@@
         \begin{enumerate}
-        	\item Add support for systems of ODEs
-        	\item Add support for PDEs
-        	\item Add all the ML operators
+        	\item Support systems of ODEs
+        	\item Support PDEs
+        	\item Add ML operators
         	\item Optimise
@@
         \begin{enumerate}
-        	\item Compare the ML operator to conventional ones
-        	\item Compare the Parareal framework with and without ML
-        	\item Measure the execution time of the Parareal (ML) framework against a traditional parallel solver
+        	\item ML operators
+        	\item Framework with and without ML
+        	\item Against another parallel solver
         \end{enumerate}
     \item Write dissertation
+        \begin{enumerate}
+        	\item Background
+        	\item Problem formulation
+        	\item Design and implementation
+        	\item Results
+        	\item Proof read
+        \end{enumerate}
 \end{enumerate}
 
+\begin{figure}[!htb]
+\begin{center}
+\ganttset{calendar week text={W\currentweek}}
+\begin{ganttchart}[
+        hgrid,
+        vgrid,
+        x unit=1mm,
+        y unit chart=5mm,
+        time slot format=isodate,
+        time slot unit=day]{2020-05-25}{2020-08-23}
+    \gantttitlecalendar{month=name, week=1} \\
+    \ganttgroup{Implement framework}{2020-05-25}{2020-07-18} \\
+    \ganttbar[bar/.style={fill=red!50}]{Support systems of ODEs}{2020-05-25}{2020-06-07} \\
+    \ganttbar[bar/.style={fill=red!50}]{Support PDEs}{2020-05-31}{2020-06-14} \\
+    \ganttbar[bar/.style={fill=red!50}]{Add ML operators}{2020-06-14}{2020-06-28} \\
+    \ganttbar[bar/.style={fill=red!50}]{Optimise}{2020-06-28}{2020-07-18} \\
+    \ganttgroup{Benchmark}{2020-07-04}{2020-07-17} \\
+    \ganttbar[bar/.style={fill=green!50}]{ML operators}{2020-07-04}{2020-07-10} \\
+    \ganttbar[bar/.style={fill=green!50}]{Framework with and without ML}{2020-07-08}{2020-07-13} \\
+    \ganttbar[bar/.style={fill=green!50}]{Against another parallel solver}{2020-07-10}{2020-07-17} \\
+    \ganttgroup{Write dissertation}{2020-05-25}{2020-08-21} \\
+    \ganttbar[bar/.style={fill=blue!50}]{Background}{2020-05-25}{2020-06-11} \\
+    \ganttbar[bar/.style={fill=blue!50}]{Problem formulation}{2020-06-12}{2020-06-28} \\
+    \ganttbar[bar/.style={fill=blue!50}]{Design and implementation}{2020-06-29}{2020-07-24} \\
+    \ganttbar[bar/.style={fill=blue!50}]{Results}{2020-07-25}{2020-08-15} \\
+    \ganttbar[bar/.style={fill=blue!50}]{Proof read}{2020-08-16}{2020-08-21}
+\end{ganttchart}
+\end{center}
+\caption{Gantt chart of the dissertation work plan}
+\label{fig:gantt}
+\end{figure}
+
 \section{Risk Analysis}
+
+We identify the main risks as:
 
@@
     		\item Parareal performance with and without machine learning
-    		\item Parareal performance against conventional parallel solver
+    		\item Framework performance against conventional parallel solvers
     	\end{enumerate}

### 2020-04-14 report/MSc-report.tex
Update on Overleaf.

--- 2020-03-26 report/MSc-report.tex
@@
 \usepackage{booktabs}
+\usepackage{pgfgantt}
+\usepackage{url}
 
@@
 Differential equations can model a wide array of dynamical systems that play important roles in science, engineering, and finance.
-Examples of these include population growth, the motion of fluids, and changes in stock prices.
+Examples of these include population growth, the motion of fluids, and stock market dynamics.
 While some simple differential equations can be solved analytically, most of them require a numerical approach \cite[p.~310]{suli2003}.
-Due to the importance of these models, their accurate and efficient numerical solution is a well researched area of numerical analysis.
-A lot of this research is focused on the parallelisation of numerical solvers to achieve better accuracy or reduce execution times.
-Traditionally, this parallelisation is applied either across the system or across time \cite{gear1988}.
+Due to the importance of these models, their accurate and efficient solution is a well researched area of numerical analysis.
+A lot of this research is focused on the parallelisation of numerical solvers to achieve higher accuracy or to reduce execution times.
+Usually, this parallelisation is applied either across the system or across time \cite{gear1988}.
+The former generally means exploiting parallelism for the evaluation or integration of the right-hand side of a differential equation at a single time point \cite[p.~1-2]{solodushkin2016}.
+While this is a straightforward approach, it only works well for high dimensional differential equations.
+In contrast, parallel-in-time methods allow for the utilisation of high levels of parallelism regardless of the dimensionality of the problem.
+
+Our project focuses on the Parareal algorithm \cite{Parareal} from the family of parallel-in-time methods.
+Given the necessary computational resources, this algorithm can solve problems faster, in terms of wall clock time, than traditional parallel solvers which can usually only take advantage of limited levels of parallelism tightly bounded by the problem size.
+This property of the algorithm makes it ideal for problems requiring real-time solutions which is why it is called "Parareal" \cite[p.~3]{staff2003}.
+
+The algorithm works by dividing the time domain up into $P$ slices where $P$ is the number of executing processes or threads.
+Each one of these $P$ time slices then represents an initial value problem (IVP) that can be solved in parallel.
+To calculate the initial values, the Parareal algorithm relies on a coarse operator $G$ that is run serially.
+Once the initial value estimates are computed, the fine operator $F$ is run in parallel across the time slices to solve the individual IVPs and provide correction terms for the subsequent initial values.
+This is an iterative process where each iteration refines the solution and the total number of iterations necessary is determined by the solution's accuracy requirements.
+
+As $G$ is executed serially, it needs to be as fast as possible to avoid creating a bottleneck that restricts parallel speed-up.
+However, it also needs to be accurate enough to keep the number of corrective iterations required at the minimum.
+Conventional coarse operators, such as those relying on numerical integration, usually trace the approximate solution using a significantly finer temporal discretisation than the extents of the $P$ time slices.
+The finer this step size is, the more accurate the initial value estimates are \cite[p.~319-320]{suli2003}.
+However, a finer step size also means more evaluations to calculate the initial values.
+Moreover, there is an additional trade-off between the sophistication of the operator and the step size.
+Higher order numerical integrators, for example, are more accurate but they are also more computationally expensive \cite[p.~325]{suli2003}.
+
+This brings us to the primary hypothesis of our project which is that a machine learning (ML) model can be trained and used as the coarse operator $G$ in a Parareal framework and it can achieve better performance, in terms of accuracy and execution time, than conventional operators.
+As there is a substantial body-of-work on using ML to solve differential equations, we begin this report by briefly introducing the most relevant and most notable pieces of research.
+We then present some of the results of our preliminary investigations into the dissertation project.
+Following that, we elaborate on our final project proposal, introduce our work plan, present our risk analysis, and outline the contents of the dissertation report.
+Lastly, we finish with a review of Nestor Sanchez' dissertation from 2018 "Fitting Large-Scale Gaussian Mixtures With Accelerated Gradient Descent" \cite{sanchez2018}.
 
@@
 
-To do...
+Parallel-in-time methods, and the Parareal algorithm in particular, have been attracting a lot of interest due to their suitability for long-time simulations and problems requiring fast solutions \cite{gurrala2015}.
+A lot of this interest has been directed to the analysis of the algorithm itself \cite{staff2003} \cite{gander2007}.
+An important property of the Parareal algorithm is that it converges to the serial solution as computed by $F$ in a maximum of $P$ iterations \cite[p.~6-7]{staff2003}.
+However, for optimal speedup, the algorithm needs to converge in much fewer iterations than $P$.
+In fact, if the computational cost of integrating over the entire domain using the operator $G$ matches the cost of integrating over a single time slice using $F$, the algorithm has to converge in fewer than $P / 2$ iterations to achieve any speedup \cite[p.~7]{staff2003}.
+
+This trade-off between the number of iterations to convergence and the computational cost of $G$ is a crucial element of the Parareal algorithm's performance characteristics.
+Operators that are faster at the same level of accuracy result in better parallel speedup due to a reduced serial bottleneck.
+Likewise, operators that are more accurate at the same speed yield better parallel speedup due to faster convergence.
+This leads us to another important property of the algorithm which is that as long as both $G$ and $F$ are convergent and stable, they can be of completely different classes of solvers and they may even solve slightly different versions of the differential equation \cite[p.~8]{staff2003}.
+Therefore, the Parareal algorithm is indifferent to how its operators work as long as they meet the above requirements and conform to the same interface.
+
+This property of the algorithm has been the focal point of most research into maximizing its speedup.
+Some of this research has tried to address the speed-accuracy dilemma of the coarse operator by relaxing its speed requirements through more efficient implementations of the Parareal algorithm using improved scheduling.
+In these event based frameworks, instead of waiting for the entire previous iteration to complete, the tasks associated with the time slices are run as soon as the necessary data is available \cite{aubanel2011} \cite{berry2012}.
+
+In contrast, others have attempted to tackle the issue directly by experimenting with different coarse operators.
+Generally, a simple Runge-Kutta method \cite{runge1895} such as Euler's method is used as $G$ and a higher order method, potentially with finer temporal discretisation, such as the Runge-Kutta 4th order (RK4) method is used as $F$ \cite[p.~1822]{gurrala2015} \cite[p.~2]{duan2016}.
+However, it has been shown that there is more to consider than just the integration method and its step size.
+In some cases, it is possible to improve the speed of the coarse operator without a significant detriment to accuracy by simplifying the differential equation (only for $G$) and thus reducing the complexity of the model it operates on \cite{duan2016}.
+Consequently, selecting and fine tuning the coarse operator is a challenging task that boils down to finding or building a differential equation solver with optimal speed and accuracy characteristics for the problem at hand.
+
+Given the success of ML in a wide range of computational fields, it is unsurprising that their application to the solution of differential equations is a popular field of research.
+Due to their prowess as universal function approximators \cite{hornik1991}, most efforts are targeted towards using artificial neural networks (ANN) to approximate the solutions of differential equations \cite{lagaris1998} \cite{lagaris2000} \cite{han2018}.
+While the majority of these works take advantage of the differentiability of neural networks and use training processes tailored explicitly to them, it has been shown that it is possible to use ML models to approximate differential equations in a strictly supervised, black-box setting as well \cite{regazzoni2019}.
+
+Perhaps the most successful approach so far to solve complex differential equations with ML is based on so called physics informed neural networks (PINN) \cite{raissi2017} \cite{tartakovsky2018} \cite{sirignano2018}.
+PINNs are architecturally standard neural networks with the values of the differential equation's independent variables ($t$ and $x$) as their input and the value of the solution at the point defined by the independent variables as their output.
+The most important feature of PINNs is their specialised loss function that penalises violations of the initial conditions, the boundary conditions, and the differential equation itself in the form of L2 terms.
+This loss function allows the network to be trained through gradient descent in a largely unsupervised fashion on collocation points sampled from the differential equation's spatiotemporal domain according to a multivariate distribution (with an additional set of points sampled from the boundaries of the spatial domain according to a different distribution).
+While PINNs are highly efficient for certain types of high dimensional partial differential equations (PDE), for many problems they are still slower than conventional solvers such as the finite element method (FEM) \cite{lu2019}.
+
+One of the main issues with PINNs is that their training can easily become prohibitively expensive for long-time simulations as the amount of data required to train them is proportional to the extent of the temporal domain.
+This is what parallel-in-time PINNs (PPINN) \cite{meng2019} address by essentially replacing the fine operator $F$ in a Parareal framework with a PINN.
+We note that while PPINNs involve both the Parareal algorithm and an ML operator, they are merely a time-parallel implementation of PINNs rather than a solution to the speed-accuracy dilemma of coarse operators.
+Therefore, PPINNs are vulnerable to the same speedup limitations imposed by potentially poor convergence or serial bottlenecks.
 
@@
 
+In order to complement our literature review and crystallise our understanding of the dissertation project, we have taken a practical approach to our initial investigations.
+As part of these investigations, we have accomplished three milestones which we briefly present below.
+
+The first one of these milestones has been the implementation of a basic 1D diffusion simulation in C.
+This simulation uses simple Dirichlet boundary conditions and Gaussian initial values.
+The spatial discretisation is done via the finite difference method and the integrator used is Euler's method.
+
+The second, somewhat more ambitious milestone has been the implementation of a simple shared-memory Parareal framework in C using POSIX threads to solve an ordinary differential equation (ODE) describing the population growth of rabbits.
+This framework employs Euler's method as $G$, RK4 as $F$, and can only handle matching coarse and fine operator step sizes.
+Although the algorithm runs for a predefined number of corrective iterations, the implementation demonstrates a significant speedup at reasonable levels of accuracy even when running on only 4 threads.
+Moreover, it showcases how the parallel solution converges to the serial fine solution as the number of iterations approaches $P$.
+
+Finally, our third and most substantial milestone has been the development of a distributed Parareal framework in Python using mpi4py \cite{dalcin2005}.
+This framework is fairly flexible in that it can handle any operator and any ODE that conform to the respective interface definitions.
+The algorithm also supports early stopping based on the absolute value of the largest update to the solution at any time step.
+In addition, the framework provides a number of Runge-Kutta method implementations and a conventional operator that takes an integrator instance and a step size value as its parameters.
+Furthermore, the framework also includes an ML accelerated operator implementation that is parameterised by a regression model, a target operator, and a step size.
+The ML operator trains the model on every new differential equation by generating a set of input-output pairs through a stochastic process using the target operator and fitting the regression model to the generated data.
+Using this approach, we have successfully trained and utilised an ML operator backed by a linear regression model as the coarse operator $G$ in our Parareal framework to solve a simple ODE.
+
+Through these milestones, we have improved our understanding of the Parareal algorithm and gained a bit of experience in applied numerical analysis.
+Moreover, we have also managed to confirm the project's feasibility on a highly simplified problem by showing that an ML accelerated coarse operator can outperform a conventional one.
+Finally, we have created a strong foundation and starting point for our dissertation project.
+
 \section{Final Proposal}
 
-\section{Workplan}
+Our final project proposal is based closely on the key steps outlined in the original proposal and it introduces only minor changes.
+The focus remains implementing a Parareal framework using an ML model as the coarse operator and solving differential equations of increasing complexity.
+Similarly, the primary metrics are still the wall time, computational cost, and accuracy of the coarse operator and the framework as a whole.
+Consequently, we formulate our research questions as
 
 \begin{enumerate}
-    \item Implement framework
-        \begin{enumerate}
-        	\item Add support for systems of ODEs
-        	\item Add support for PDEs
-        	\item Add all the ML operators
-        	\item Optimise
-        \end{enumerate}
-    \item Benchmark
-        \begin{enumerate}
-        	\item Compare the ML operator to conventional ones
-        	\item Compare the Parareal framework with and without ML
-        	\item Measure the execution time of the Parareal (ML) framework against a traditional parallel solver
-        \end{enumerate}
-    \item Write dissertation
+    \item how does an ML accelerated coarse operator compare to traditional coarse operators in terms of wall time and accuracy
+    \item is an ML accelerated Parareal framework faster than a conventional Parareal framework given the same accuracy requirements
+    \item is an ML accelerated Parareal framework faster than other conventional parallel solvers given the same accuracy requirements and using the same number of parallel processes
 \end{enumerate}
 
+To answer our first research question, we propose training and evaluating a number of different regression models to replace the coarse operator.
+The models we are interested in trying include linear regression, decision tree ensembles such as boosted trees and random forests, and ANNs.
+All of these models can be treated as black boxes in terms of both training and evaluation.
+Their output is the estimate of the solution of the differential equation at the next time step (as determined by the step size of the operator) while their inputs are some combination or subset of the values of the independent variables, the derivatives of the solution, and the estimate of the solution at the provided values of the independent variables.
+These target outputs are supplied by a conventional operator which may be the intended fine operator of the framework or an entirely unrelated one.
+
+In addition, we also intend to try sequence models such as recurrent neural networks (RNN) to capture temporal patterns in the differential equations' solutions.
+We hypothesize that this increases the robustness of the learned models and aids the approximation of periodic solutions.
+However, this requires the capturing of inputs from possibly multiple previous time steps.
+Therefore, we propose a more general training and evaluation scheme where the input is a sequence.
+The training of the non-recurrent models would be, thus, merely a special case of this scheme where the input sequence contains only one element.
+This allows for the implementation of a single ML accelerated operator that is agnostic of the model.
+
+Finally, we also intend to train and evaluate a separate class of ML accelerated operators backed by PINNs.
+As the training process of PINNs is significantly different, they are not interchangeable with other models.
+However, due to the fact that this training process is largely unsupervised and mesh-free, PINNs are likely the best option for higher dimensional PDEs.
+Therefore, we propose implementing an additional, PINN based ML accelerated operator to be used as the coarse operator of our Parareal framework.
+This differs from \cite{meng2019} in that employing a PINN as the coarse operator, instead of the fine one, allows us to increase the step size and thus minimise the number of inferences without compromising the fine trajectory of our solution (as the coarse operator only needs to provide the estimates of the initial conditions of the time slices and not the fine solution).
+Consequently, this enables us to experiment with more complex network architectures in favour of higher accuracy at a lower cost to wall time.
+
+We propose implementing our operators in Python using sklearn \cite{pedregosa2011} for linear regression and the decision tree based models and either Tensorflow \cite{mart2015} and Keras \cite{chollet2015} or PyTorch \cite{paszke2019} for the neural network models.
+As for our benchmarks, we plan to compare the execution time and accuracy of the ML accelerated operators to conventional solvers provided by scipy \cite{scipy2020}, FEniCS \cite{fenics2015}, and FiPy \cite{fipy2009} (depending on the type of the differential equation).
+Additionally, we are also interested in the training times of the ML operators using different models.
+For differential equations without an analytic solution, we base our accuracy comparisons on the solutions of higher order, small step-size solvers.
+
+To answer our second question, we propose extending the Python Parareal framework implemented as part of our preliminary investigations.
+As most of the computationally expensive operations are delegated to third party libraries with C bindings, we expect the performance cost of using Python to be minimal.
+In case the overhead proves to be more significant than expected, we can further optimize our Parareal implementation using static typing with Cython \cite{behnel2010}.
+
+For the comparison between the ML accelerated and the normal versions of our Parareal framework, we are interested in the wall time of the solution using the same number of processes and the same accuracy requirements.
+To ensure our analysis is comprehensive enough, we propose performing comparisons across different combinations of process counts and accuracy levels.
+We intend to time the ML accelerated version both including and excluding the training of the models.
+Furthermore, we also aim to consider the amortized cost of model training through transfer learning.
+We plan to do so by quantifying how well a model trained on one IVP performs on a different IVP expressed by the same differential equation.
+We propose using the same approach to answer the third research question as well where we compare our ML accelerated Parareal framework to conventional parallel ODE and PDE solvers provided by FEniCS.
+
+Finally, as for the differential equations to benchmark our operators and framework on, we set forth three different problems of increasing complexity.
+The first one is the Lotka-Volterra model \cite{lotka1926} \cite{volterra1927} which is a system of two ODEs expressing the dynamics of the populations of predators and their preys.
+This requires the extension of our initial framework to handle systems of differential equations which is the same as handling differential equations with vector valued solutions.
+The second differential equation we propose solving is the two dimensional heat equation.
+This requires the extension of the framework to handle PDEs, that is differential equations of multiple independent variables.
+Lastly, the third problem we consider is the Black-Scholes equation \cite{black1973} which is a high dimensional PDE that can be computationally challenging to solve using traditional mesh-based approaches.
+
+\section{Work Plan}
+
+In this section, we outline our work plan for the dissertation period beginning on 25th May and ending on 21st August. The units of work listed below are based on our final project proposal.
+
+\begin{enumerate}
+    \item Implementation
+        \begin{enumerate}
+        	\item Testing
+        	\item Systems of ODEs
+        	\item PDEs
+        	\item ML operators
+        	\item Optimisation
+        \end{enumerate}
+    \item Benchmarking
+        \begin{enumerate}
+        	\item Only ML operators
+        	\item Framework with and without ML
+        	\item Against another parallel solver
+        \end{enumerate}
+    \item Writing
+        \begin{enumerate}
+            \item Engineering log
+            \item Phase 1
+            \item Phase 2
+        	\item Proof reading
+        \end{enumerate}
+\end{enumerate}
+
+The three main tasks of the dissertation we identify are the implementation of the software framework, the benchmarking of its performance, and the write up of the dissertation report.
+Each one of these high level tasks are divided up into smaller sub-tasks.
+For the implementation task, these sub-tasks are software testing, the extension of the Parareal framework to handle systems of ODEs and individual PDEs, the implementation of the two final ML accelerated operators discussed in the project proposal (one for PINNs and one for all the other models), and code optimisation.
+As for the benchmarking task, the sub-tasks are the comparison of the ML accelerated operator to conventional operators, the comparison of the Parareal framework using an ML accelerated coarse operator to the same framework using a conventional coarse operator, and the comparison of the ML accelerated Parareal framework to other parallel solvers.
+Finally, the last main task, the writing of the dissertation consists of four smaller units of work.
+These are the engineering log to keep during the initial development period, the first phase for writing up the problem formulation and the design and implementation sections of the dissertation, the second phase for writing up the rest of the dissertation such as the literature review and the benchmark results, and lastly, the proof reading of the report.
+
+\begin{figure}[!htb]
+\begin{center}
+\ganttset{calendar week text={W\currentweek}}
+\begin{ganttchart}[
+        hgrid,
+        vgrid,
+        x unit=1mm,
+        y unit chart=5mm,
+        time slot format=isodate,
+        time slot unit=day]{2020-05-25}{2020-08-23}
+    \gantttitlecalendar{month=name, week=1} \\
+    \ganttgroup{Implementation}{2020-05-25}{2020-07-18} \\
+    \ganttbar[bar/.style={fill=red!50}]{Testing}{2020-05-25}{2020-06-28} \\
+    \ganttbar[bar/.style={fill=red!50}]{Systems of ODEs}{2020-05-25}{2020-06-07} \\
+    \ganttbar[bar/.style={fill=red!50}]{PDEs}{2020-05-31}{2020-06-14} \\
+    \ganttbar[bar/.style={fill=red!50}]{ML operators}{2020-06-15}{2020-06-28} \\
+    \ganttbar[bar/.style={fill=red!50}]{Optimisation}{2020-06-29}{2020-07-18} \\
+    \ganttgroup{Benchmarking}{2020-07-04}{2020-07-17} \\
+    \ganttbar[bar/.style={fill=green!50}]{Only ML operators}{2020-07-04}{2020-07-10} \\
+    \ganttbar[bar/.style={fill=green!50}]{Framework with and without ML}{2020-07-08}{2020-07-13} \\
+    \ganttbar[bar/.style={fill=green!50}]{Against another parallel solver}{2020-07-10}{2020-07-17} \\
+    \ganttgroup{Writing}{2020-05-25}{2020-08-21} \\
+    \ganttbar[bar/.style={fill=blue!50}]{Engineering log}{2020-05-25}{2020-06-28} \\
+    \ganttbar[bar/.style={fill=blue!50}]{Phase 1}{2020-06-29}{2020-08-02} \\
+    \ganttbar[bar/.style={fill=blue!50}]{Phase 2}{2020-08-03}{2020-08-20} \\
+    \ganttbar[bar/.style={fill=blue!50}]{Proof read}{2020-08-16}{2020-08-21}
+\end{ganttchart}
+\end{center}
+\caption{Gantt chart of the dissertation work plan}
+\label{fig:gantt}
+\end{figure}
+
+Figure \ref{fig:gantt} contains a Gantt chart displaying the timeline of the dissertation project and the different tasks it encompasses.
+Although the times are just rough and somewhat optimistic estimates, the chart sufficiently visualises the overlaps between certain tasks.
+For example, it clearly captures the non-linear nature of optimisation and benchmarking as they feed into each other.
+
 \section{Risk Analysis}
+
+We identify the main risks of the dissertation project as
 
@@
 \section{Outline of the Dissertation Report}
+
+Based on our final project proposal and work plan, we set forth the following structure of contents for the dissertation report.
 

### 2020-04-17 report/MSc-report.tex
Update on Overleaf.

@@
 \usepackage{url}
+\usepackage{hyperref}
+\usepackage[thinc]{esdiff}
+\usepackage{tabularx}
+
+\newcolumntype{Y}{>{\centering\arraybackslash}X}
 
@@
 
-Our project focuses on the Parareal algorithm \cite{Parareal} from the family of parallel-in-time methods.
+Our project focuses on the Parareal algorithm \cite{staff2003} from the family of parallel-in-time methods.
 Given the necessary computational resources, this algorithm can solve problems faster, in terms of wall clock time, than traditional parallel solvers which can usually only take advantage of limited levels of parallelism tightly bounded by the problem size.
@@
 
-The algorithm works by dividing the time domain up into $P$ slices where $P$ is the number of executing processes or threads.
+The Parareal algorithm solves time-dependent differential equations by splitting them up along the time domain and assigning a sub-problem to each processor.
+It first computes the initial values for each sub-problem serially using a coarse operator.
+It then solves each of these problems based on the initial values in parallel using a fine operator and calculates a correction to the initial values.
+This iterative process goes on until a sufficient level of accuracy is reached.
+As the coarse operator is executed serially at every iteration, it has to be fast but it also needs to be accurate enough for rapid convergence.
+
+This brings us to the primary hypothesis of our project which is that a machine learning (ML) model can be trained and used as the coarse operator in a Parareal framework and it can achieve better performance, in terms of accuracy and execution time, than conventional operators.
+As there is a substantial body-of-work on using ML to solve differential equations, we begin this report by briefly introducing the most relevant and most notable pieces of research.
+We then present some of the results of our preliminary investigations into the dissertation project.
+Following that, we elaborate on our final project proposal, introduce our work plan, present our risk analysis, and outline the contents of the dissertation report.
+Lastly, we finish with a review of Nestor Sanchez' dissertation from 2018 "Fitting Large-Scale Gaussian Mixtures With Accelerated Gradient Descent" \cite{sanchez2018}.
+
+\section{Background and Literature Review}
+
+The Parareal algorithm works by dividing the time domain up into $P$ slices where $P$ is the number of executing processes or threads.
 Each one of these $P$ time slices then represents an initial value problem (IVP) that can be solved in parallel.
-To calculate the initial values, the Parareal algorithm relies on a coarse operator $G$ that is run serially.
-Once the initial value estimates are computed, the fine operator $F$ is run in parallel across the time slices to solve the individual IVPs and provide correction terms for the subsequent initial values.
+To calculate the initial values, the Parareal algorithm relies on a coarse operator, $G$, that is run serially.
+Once the initial value estimates are computed, the fine operator, $F$, is run in parallel across the time slices to solve the individual IVPs and provide correction terms for the subsequent initial values.
 This is an iterative process where each iteration refines the solution and the total number of iterations necessary is determined by the solution's accuracy requirements.
@@
 
-This brings us to the primary hypothesis of our project which is that a machine learning (ML) model can be trained and used as the coarse operator $G$ in a Parareal framework and it can achieve better performance, in terms of accuracy and execution time, than conventional operators.
-As there is a substantial body-of-work on using ML to solve differential equations, we begin this report by briefly introducing the most relevant and most notable pieces of research.
-We then present some of the results of our preliminary investigations into the dissertation project.
-Following that, we elaborate on our final project proposal, introduce our work plan, present our risk analysis, and outline the contents of the dissertation report.
-Lastly, we finish with a review of Nestor Sanchez' dissertation from 2018 "Fitting Large-Scale Gaussian Mixtures With Accelerated Gradient Descent" \cite{sanchez2018}.
-
-\section{Background and Literature Review}
-
-Parallel-in-time methods, and the Parareal algorithm in particular, have been attracting a lot of interest due to their suitability for long-time simulations and problems requiring fast solutions \cite{gurrala2015}.
+The Parareal algorithm has been attracting a lot of interest due to its suitability for long-time simulations and problems requiring fast solutions \cite{gurrala2015}.
 A lot of this interest has been directed to the analysis of the algorithm itself \cite{staff2003} \cite{gander2007}.
-An important property of the Parareal algorithm is that it converges to the serial solution as computed by $F$ in a maximum of $P$ iterations \cite[p.~6-7]{staff2003}.
-However, for optimal speedup, the algorithm needs to converge in much fewer iterations than $P$.
+An important property of the algorithm is that it converges to the serial solution as computed by $F$ in a maximum of $P$ iterations \cite[p.~6-7]{staff2003}.
+However, for optimal speedup, it needs to converge in much fewer iterations than $P$.
 In fact, if the computational cost of integrating over the entire domain using the operator $G$ matches the cost of integrating over a single time slice using $F$, the algorithm has to converge in fewer than $P / 2$ iterations to achieve any speedup \cite[p.~7]{staff2003}.
@@
 In contrast, others have attempted to tackle the issue directly by experimenting with different coarse operators.
-Generally, a simple Runge-Kutta method \cite{runge1895} such as Euler's method is used as $G$ and a higher order method, potentially with finer temporal discretisation, such as the Runge-Kutta 4th order (RK4) method is used as $F$ \cite[p.~1822]{gurrala2015} \cite[p.~2]{duan2016}.
+Generally, a simple Runge-Kutta method such as the forward Euler method is used as $G$ and a higher order method, potentially with finer temporal discretisation, such as the Runge-Kutta 4th order (RK4) method is used as $F$ \cite[p.~1822]{gurrala2015} \cite[p.~2]{duan2016}.
 However, it has been shown that there is more to consider than just the integration method and its step size.
@@
 Given the success of ML in a wide range of computational fields, it is unsurprising that their application to the solution of differential equations is a popular field of research.
+The most attractive prospect of using ML to this end is that a trained model can directly represent the solution without the need for any discretisation (which can be intractable for high dimensional problems).
 Due to their prowess as universal function approximators \cite{hornik1991}, most efforts are targeted towards using artificial neural networks (ANN) to approximate the solutions of differential equations \cite{lagaris1998} \cite{lagaris2000} \cite{han2018}.
-While the majority of these works take advantage of the differentiability of neural networks and use training processes tailored explicitly to them, it has been shown that it is possible to use ML models to approximate differential equations in a strictly supervised, black-box setting as well \cite{regazzoni2019}.
+This is also partly attributable to the fact that ANNs are differentiable which allows for more efficient training approaches.
+While the majority of recent works take advantage of the differentiability of neural networks and use training processes tailored explicitly to them, it has been shown that it is possible to use ML models to approximate the solutions of differential equations in a strictly supervised, black-box setting as well \cite{regazzoni2019}.
 
-Perhaps the most successful approach so far to solve complex differential equations with ML is based on so called physics informed neural networks (PINN) \cite{raissi2017} \cite{tartakovsky2018} \cite{sirignano2018}.
-PINNs are architecturally standard neural networks with the values of the differential equation's independent variables ($t$ and $x$) as their input and the value of the solution at the point defined by the independent variables as their output.
-The most important feature of PINNs is their specialised loss function that penalises violations of the initial conditions, the boundary conditions, and the differential equation itself in the form of L2 terms.
-This loss function allows the network to be trained through gradient descent in a largely unsupervised fashion on collocation points sampled from the differential equation's spatiotemporal domain according to a multivariate distribution (with an additional set of points sampled from the boundaries of the spatial domain according to a different distribution).
-While PINNs are highly efficient for certain types of high dimensional partial differential equations (PDE), for many problems they are still slower than conventional solvers such as the finite element method (FEM) \cite{lu2019}.
+Perhaps the most successful approach so far to solve complex differential equations with ML is based on physics informed neural networks (PINN) \cite{raissi2017} \cite{tartakovsky2018} \cite{sirignano2018}.
+PINNs approximate the function that solves the differential equation, therefore their inputs are the values of the differential equation's independent variables ($t$ and $x$) and their output is the value of the solution at the point defined by the inputs.
+In terms of architecture, PINNs are flexible and can be implemented by simple ANNs, convolutional neural networks, and even recurrent neural networks (RNN).
+The most important feature of PINNs is their specialised loss function that penalises violations of the initial conditions, the boundary conditions, and the differential equation itself in the form of mean squared error terms.
+By minimising this loss function, for example through gradient descent, the network can be trained in a largely unsupervised fashion on collocation points sampled from the differential equation's spatiotemporal domain according to a multivariate distribution (with an additional set of points sampled from the boundaries of the domain according to a different distribution).
+Although training a PINN can be very costly, especially on differential equations with a large domain, a trained model may perform more than an order of magnitude faster than conventional solvers at comparable levels of accuracy \cite{misyris2019}.
+However, while PINNs are fully mesh-free and very efficient for high dimensional partial differential equations (PDE), for some problems they are still slower than conventional solvers such as the finite element method \cite{lu2019}.
 
@@
 
-The first one of these milestones has been the implementation of a basic 1D diffusion simulation in C.
+\begin{equation}
+    \diffp{u}{t} = d \diffp{u}{{x^2}}
+\label{equation:1d_diff}
+\end{equation}
+
+The first one of these milestones has been the implementation of a basic one dimensional diffusion simulation in C defined by the model in Equation \ref{equation:1d_diff} where $u$ is the unknown function, $x$ is the spatial coordinate, $t$ is time, and $d$ is the diffusion coefficient.
 This simulation uses simple Dirichlet boundary conditions and Gaussian initial values.
-The spatial discretisation is done via the finite difference method and the integrator used is Euler's method.
+The spatial discretisation is done via the finite difference method and the integrator used is the forward Euler method.
+The implementation can be found at \href{https://git.ecdf.ed.ac.uk/msc-19-20/s1984842/blob/master/code/c/src/main.c}{\texttt{code/c/src/main.c}} in the project repository.
+
+\begin{equation}
+    \diff{R}{t} = \alpha R
+\label{equation:rabbit_pop}
+\end{equation}
 
 The second, somewhat more ambitious milestone has been the implementation of a simple shared-memory Parareal framework in C using POSIX threads to solve an ordinary differential equation (ODE) describing the population growth of rabbits.
-This framework employs Euler's method as $G$, RK4 as $F$, and can only handle matching coarse and fine operator step sizes.
+This ODE is expressed by Equation \ref{equation:rabbit_pop} where $R$ is the number of rabbits and $\alpha$ is the population growth rate.
+The framework employs the forward Euler method as $G$ and RK4 as $F$ but it can only handle matching coarse and fine operator step sizes.
 Although the algorithm runs for a predefined number of corrective iterations, the implementation demonstrates a significant speedup at reasonable levels of accuracy even when running on only 4 threads.
 Moreover, it showcases how the parallel solution converges to the serial fine solution as the number of iterations approaches $P$.
+This implementation can be found in the same file as the diffusion simulation.
 
@@
 The ML operator trains the model on every new differential equation by generating a set of input-output pairs through a stochastic process using the target operator and fitting the regression model to the generated data.
+The data generation involves repeatedly integrating over the differential equation by the target operator and adding a $0$-mean Gaussian noise to the estimate at every time step to use as the initial value for the next step and thus perturbate the trajectory of the target operator's solution.
 Using this approach, we have successfully trained and utilised an ML operator backed by a linear regression model as the coarse operator $G$ in our Parareal framework to solve a simple ODE.
+The implementation of this framework and our experiment can be found in the \href{https://git.ecdf.ed.ac.uk/msc-19-20/s1984842/blob/master/code/python/}{\texttt{code/python}} directory of the project repository
 
-Through these milestones, we have improved our understanding of the Parareal algorithm and gained a bit of experience in applied numerical analysis.
+\begin{table}[!htb]
+\begin{center}
+\begin{tabularx}{\textwidth}{Y|Y|Y|}
+\multicolumn{1}{c}{} & \multicolumn{2}{c}{Operator}\\
+\cmidrule(l){2-3}
+& Explicit midpoint & Linear regression\\
+\midrule
+Run time (s) & 0.233 & \textbf{0.017}\\
+\bottomrule
+\end{tabularx}
+\end{center}
+\caption{The average execution times of different coarse operators on the rabbit population problem}
+\label{tab:coarse_times}
+\end{table}
+
+Table \ref{tab:coarse_times} shows the average execution times, based on $10$ runs, of two different coarse operators as they integrate over the entire time domain of the rabbit population problem defined by Equation \ref{equation:rabbit_pop} using the values $R_0 = 1000$, $\alpha = 0.0001$, and $t \in [0, 50000]$.
+The first operator is a conventional one using the explicit midpoint integration method and a step size of $0.25$ while the second operator is an ML accelerated one with a step size of $100$ backed by an sklearn \cite{pedregosa2011} linear regression model trained on data labelled by the first operator.
+As seen in the table, the ML accelerated operator solves the differential equation almost $14$ times faster with an average deviation of less than $10^{-6}$ from the conventional operator's solution.
+This performance difference is due to the fact that the ML operator's step size is $400$ times greater and therefore the solution is evaluated $400$ times fewer while still maintaining the same level of accuracy.
+
+\begin{table}[!htb]
+\begin{center}
+\begin{tabularx}{\textwidth}{Y|Y|Y|Y|Y|}
+\multicolumn{1}{c}{} & \multicolumn{4}{c}{Solver}\\
+\cmidrule(l){2-5}
+& Serial & Parareal & ML Parareal & ML Parareal with training\\
+\midrule
+Run time (s) & 8.387 & 6.082 & \textbf{5.080} & 8.020\\
+\bottomrule
+\end{tabularx}
+\end{center}
+\caption{The average execution times of different solvers on the rabbit population problem}
+\label{tab:parareal_times}
+\end{table}
+
+Table \ref{tab:parareal_times} shows the average execution times, again across $10$ runs, of different solvers on the same rabbit population problem.
+The first, serial solver is a conventional operator using the RK4 integration method and a step size of $0.01$.
+The others are all Parareal solvers running on $4$ processes using the first solver as their fine operator and $0.001$ as their update threshold (which means that the algorithm finishes once the absolute value of the largest update to the solution at any point of the fine operator's temporal discretisation is below this threshold).
+The first Parareal solver uses the explicit midpoint operator from Table \ref{tab:coarse_times} as its coarse operator while the second and third Parareal solvers use the linear regression operator.
+Both the conventional and the ML accelerated Parareal solvers converge after just one corrective iteration and thus are able to achieve a significant speedup compared to the serial solver while producing a solution identical to the serial one to $6$ and $5$ significant figures respectively.
+While the ML accelerated operator is slightly less accurate, it still converges within the same number of iterations given our accuracy threshold and it actually solves the differential equation a whole second faster.
+In contrast, when taking the model training into account, the ML accelerated Parareal solver is barely  faster than the serial solver.
+However, the training time could potentially be significantly reduced for the given accuracy requirements by decreasing the volume of training data generated.
+
+Through these investigations and experiments presented above, we have improved our understanding of the Parareal algorithm and gained a little experience in applied numerical analysis.
 Moreover, we have also managed to confirm the project's feasibility on a highly simplified problem by showing that an ML accelerated coarse operator can outperform a conventional one.
@@
 
-Our final project proposal is based closely on the key steps outlined in the original proposal and it introduces only minor changes.
-The focus remains implementing a Parareal framework using an ML model as the coarse operator and solving differential equations of increasing complexity.
-Similarly, the primary metrics are still the wall time, computational cost, and accuracy of the coarse operator and the framework as a whole.
+The main focus of our project is implementing a Parareal framework using an ML model as the coarse operator and solving differential equations of increasing complexity.
+The primary metrics we are interested in are the wall time, computational cost, and accuracy of the coarse operator and the framework as a whole.
 Consequently, we formulate our research questions as
@@
 \begin{enumerate}
-    \item how does an ML accelerated coarse operator compare to traditional coarse operators in terms of wall time and accuracy
-    \item is an ML accelerated Parareal framework faster than a conventional Parareal framework given the same accuracy requirements
-    \item is an ML accelerated Parareal framework faster than other conventional parallel solvers given the same accuracy requirements and using the same number of parallel processes
+    \item How do ML accelerated coarse operators compare to traditional coarse operators in terms of accuracy, execution time, and training time?
+    \item Is an ML accelerated Parareal framework faster, both excluding and including model training time, than a conventional Parareal framework given the same accuracy requirements?
+    \item Is an ML accelerated Parareal framework faster, both excluding and including model training time, than other conventional parallel solvers given the same accuracy requirements and using the same number of parallel processes?
 \end{enumerate}
@@
 
-In addition, we also intend to try sequence models such as recurrent neural networks (RNN) to capture temporal patterns in the differential equations' solutions.
+In addition, we also intend to try sequence models such as RNNs to capture temporal patterns in the differential equations' solutions.
 We hypothesize that this increases the robustness of the learned models and aids the approximation of periodic solutions.
@@
 Therefore, we propose a more general training and evaluation scheme where the input is a sequence.
-The training of the non-recurrent models would be, thus, merely a special case of this scheme where the input sequence contains only one element.
+The training of the non-recurrent models is, thus, merely a special case of this scheme where the input sequence contains only one element.
 This allows for the implementation of a single ML accelerated operator that is agnostic of the model.
@@
 
-We propose implementing our operators in Python using sklearn \cite{pedregosa2011} for linear regression and the decision tree based models and either Tensorflow \cite{mart2015} and Keras \cite{chollet2015} or PyTorch \cite{paszke2019} for the neural network models.
+While we are optimistic about machine learning accelerated operators with a greater step size outperforming conventional operators in terms of execution time at similar levels of accuracy, we are conscious of the overhead introduced by training data generation and model fitting.
+Therefore, we propose measuring the average training time for all the models tested.
+Moreover, due to the inherently stochastic nature of the training processes, we propose basing all our results on the means and standard deviations of measurements across a high number of runs and using fixed random seeds to facilitate reproducibility.
+Even though the combined training time and execution time of a machine learning operator is more than likely to significantly exceed the execution time of a conventional operator, this may be mitigated in a Parareal framework where a model is trained only once but used for inference potentially several times.
+Therefore, by keeping track of the training times of the models, we can establish a more complete picture of their performance potential as coarse operators in a Parareal setting.
+
+We propose implementing our operators in Python using sklearn for linear regression and the decision tree based models and either Tensorflow \cite{mart2015} and Keras \cite{chollet2015} or PyTorch \cite{paszke2019} for the neural network models.
 As for our benchmarks, we plan to compare the execution time and accuracy of the ML accelerated operators to conventional solvers provided by scipy \cite{scipy2020}, FEniCS \cite{fenics2015}, and FiPy \cite{fipy2009} (depending on the type of the differential equation).
-Additionally, we are also interested in the training times of the ML operators using different models.
-For differential equations without an analytic solution, we base our accuracy comparisons on the solutions of higher order, small step-size solvers.
 
@@
 We intend to time the ML accelerated version both including and excluding the training of the models.
-Furthermore, we also aim to consider the amortized cost of model training through transfer learning.
+Furthermore, as a stretch goal, we also aim to consider the amortized cost of model training through transfer learning.
 We plan to do so by quantifying how well a model trained on one IVP performs on a different IVP expressed by the same differential equation.
 We propose using the same approach to answer the third research question as well where we compare our ML accelerated Parareal framework to conventional parallel ODE and PDE solvers provided by FEniCS.
+For differential equations without an analytic solution, we plan to base our accuracy comparisons on the solutions of higher-order, smaller-step size solvers.
 
 Finally, as for the differential equations to benchmark our operators and framework on, we set forth three different problems of increasing complexity.
-The first one is the Lotka-Volterra model \cite{lotka1926} \cite{volterra1927} which is a system of two ODEs expressing the dynamics of the populations of predators and their preys.
+The first one is the Lotka-Volterra model \cite{lotka1926} which is a system of two ODEs expressing the dynamics of the populations of predators and their preys.
 This requires the extension of our initial framework to handle systems of differential equations which is the same as handling differential equations with vector valued solutions.
@@
 
-In this section, we outline our work plan for the dissertation period beginning on 25th May and ending on 21st August. The units of work listed below are based on our final project proposal.
+In this section, we outline our work plan for the dissertation period beginning on 25th May and ending on 21st August.
+The three main tasks of the dissertation we identify, based on our final project proposal, are the implementation of the software framework, the benchmarking of its performance, and the write up of the dissertation report.
+Each one of these high level tasks are divided up into smaller sub-tasks.
 
@@
         	\item ML operators
-        	\item Optimisation
+        	\item Performance optimisation
         \end{enumerate}
@@
 
-The three main tasks of the dissertation we identify are the implementation of the software framework, the benchmarking of its performance, and the write up of the dissertation report.
-Each one of these high level tasks are divided up into smaller sub-tasks.
-For the implementation task, these sub-tasks are software testing, the extension of the Parareal framework to handle systems of ODEs and individual PDEs, the implementation of the two final ML accelerated operators discussed in the project proposal (one for PINNs and one for all the other models), and code optimisation.
+For the implementation task, these sub-tasks are software testing, the extension of the Parareal framework to handle systems of ODEs and individual PDEs, the implementation of the two final ML accelerated operators discussed in the project proposal (one for PINNs and one for all the other models), and code optimisation along with hyperparameter tuning.
 As for the benchmarking task, the sub-tasks are the comparison of the ML accelerated operator to conventional operators, the comparison of the Parareal framework using an ML accelerated coarse operator to the same framework using a conventional coarse operator, and the comparison of the ML accelerated Parareal framework to other parallel solvers.
@@
     \ganttbar[bar/.style={fill=red!50}]{ML operators}{2020-06-15}{2020-06-28} \\
-    \ganttbar[bar/.style={fill=red!50}]{Optimisation}{2020-06-29}{2020-07-18} \\
+    \ganttbar[bar/.style={fill=red!50}]{Performance otimisation}{2020-06-29}{2020-07-18} \\
     \ganttgroup{Benchmarking}{2020-07-04}{2020-07-17} \\
@@
     \ganttbar[bar/.style={fill=blue!50}]{Phase 2}{2020-08-03}{2020-08-20} \\
-    \ganttbar[bar/.style={fill=blue!50}]{Proof read}{2020-08-16}{2020-08-21}
+    \ganttbar[bar/.style={fill=blue!50}]{Proof reading}{2020-08-16}{2020-08-21}
 \end{ganttchart}
@@
 Although the times are just rough and somewhat optimistic estimates, the chart sufficiently visualises the overlaps between certain tasks.
-For example, it clearly captures the non-linear nature of optimisation and benchmarking as they feed into each other.
+For example, it clearly captures the non-linear nature of optimisation and benchmarking as they feed back into each other.
+
+As for hardware requirements, our plan is to use the author's personal computer for most of the implementation phase.
+This computer has $6$ physical cores which is sufficient for development and testing purposes.
+However for optimisation and benchmarking, we plan to use high performance computing services.
+Due to the generally low number of processors needed for our benchmarks, we estimate our required budget to be about $2000$ to $3000$ CPU core hours on Cirrus.
+Additionally, if the new Cirrus GPUs are installed by late June, we also plan to take advantage of them for our neural network based models.
+For this, our required budget estimate is $50$ GPU hours.
 
@@
 
-We identify the main risks of the dissertation project as
+We identify the following as the main risks of the dissertation project.
 
 \begin{itemize}
-    \item Failure to implement a single framework to handle both ODEs and PDEs
-    \item Failure to fit a regression model to complex differential equations
-    \item Failure to utilise transfer learning
+    \item Failure to implement a single framework to handle both ODEs and PDEs in time
+    \item Some of the model training approaches may not work on complex problems
+    \item Erratic model behaviour due to the stochastic nature of the training processes
+    \item Transfer learning may not work
     \item Unavailability of computing resources
@@
 \end{itemize}
+
+The first risk we identify is the author's failure to implement a Parareal framework that can handle ODEs, systems of ODEs, and PDEs in time for our benchmarks that need to be performed early enough to leave sufficient time for writing up the results.
+Failure to do this would likely restrict our benchmarks to ODEs only which are not representative of more complex real life problems.
+However, this risk is mitigated by the fact that we already have a strong base in the form of our Parareal framework implementation that can handle any ODE.
+Moreover, we have a good idea of how to extend this framework to enable systems of ODEs and PDEs as well.
+
+The second major risk is that some of the models or training processes we intend to try may simply not work.
+While this is a high probability risk, it is rather low impact due to the fact that we plan to try multiple approaches, some of which have already been shown to work on several problems (such as PINNs).
+Additionally, we have already succeeded in training an ML accelerated operator that can outperform some conventional solvers on a very simple ODE.
+
+The third risk is erratic model behaviour due to the stochastic nature of training.
+As both our proposed supervised training approach and the unsupervised training method of PINNs have elements of randomness (and the weights of neural networks are initialised randomly as well), the models may behave fairly differently between runs.
+To address this, we plan to base our results on high numbers of runs.
+For the sake of reproducibility, we also intend to use fixed random seeds in our benchmarks.
+
+The next main risk we identify is that transfer learning may not work.
+This means that a model trained on a smaller subset of a differential equation's domain may not be able to infer the solution well enough outside of that subset.
+Alternatively, a model trained on a differential equation using certain initial value and boundary conditions may not do well on the same equation with different conditions.
+This is another high probability risk, but utilising transfer learning is only a secondary goal of our project and we believe that our results would be valuable even without it.
+
+The fifth risk is the unavailability of computing resources.
+This includes multiple possibilities.
+Cirrus may be less available towards the end of the dissertation period when most students run their experiments.
+It may not get the new GPUs in time for our benchmarks.
+Finally, it may become completely unavailable due to unforeseen circumstances.
+Therefore, it is important that we have a fallback plan in the form of alternative computing resources.
+In case Cirrus becomes unavailable for any reason, we can fall back to using Archer.
+If Archer becomes unavailable as well, we can use the author's computer which should be sufficient for smaller scale benchmarks.
+As for the GPUs, if Cirrus does not get the upgrade in time, we may try to find an alternative GPU computing service or just use CPUs for our neural networks instead.
+
+Finally, the last risk we identify is potential disruptions due to Covid-19.
+A prolonged pandemic and lockdown may have serious economical implications in the United Kingdom that may indirectly affect the dissertation project.
+While these effects are hard to predict or prevent, we intend to maintain our focus on the project as our main priority and carry on with our regular remote meetings with the project supervisors.
 
@@
     	\end{enumerate}
-	\item Problem formulation
+	\item Problem statement
     	\begin{enumerate}
@@
 
-Fitting Large-Scale Gaussian Mixtures With Accelerated Gradient Descent, Nestor Sanchez, 2018.
+In this section, we briefly review the 2018 dissertation "Fitting Large-Scale Gaussian Mixtures With Accelerated Gradient Descent" \cite{sanchez2018} by Nestor Sanchez Guadarrama.
+
+The dissertation presents a new approach to optimise a class of clustering algorithms called Gaussian mixture models (GMM).
+GMMs are made up of a fixed number of Gaussian distributions that best describe clusters in a data set.
+To fit the parameters (mean and covariance) of these distributions to the data, usually the expectation maximisation (EM) algorithm is used.
+However, as EM requires the entire data set to be available all at once, it is unsuitable for high volume or high velocity data.
+To enable the usage of GMMs for stream processing and massive data sets, Sanchez describes fitting the models using a stochastic gradient descent algorithm (SGD) on a reformulation of the GMM optimisation problem.
+This reformulation enforces the symmetry and positive-definiteness constraints on the covariance parameters of the Gaussian distributions and thus enables mini-batch optimisation through gradient descent.
+In addition, he also presents a logarithmic barrier regularisation term to avoid covariance singularities (which are formed by the optimiser minimising the covariance to a null-matrix).
+
+After explaining the theoretical foundations of the dissertation in the second and third sections, Sanchez goes on to present the design and details of his distributed implementation of the SGD optimisable GMM using Scala and Spark.
+He briefly introduces the main components of his library which include, among others, two regularisation schemes and five variations of the SGD algorithm.
+He then demonstrates the usage of the library through short examples and code snippets.
+
+In the fifth section, Sanchez begins by introducing his automated test setup and describes how the synthetic test data is generated using the parameters read from a configuration file.
+He then presents the results of some qualitative behaviour tests according to which (Nesterov) momentum accelerated SGD with logarithmic barrier regularisation performs the best.
+This is followed by the results of some hyperparameter tuning of the optimisers and regularisers which confirm that the effect of logarithmic barrier regularisation is much less detrimental to model quality than that of conjugate prior regularisation.
+Finally, the dissertation is concluded with the general performance and data streaming test results.
+The serial performance test results again confirm (Nesterov) momentum accelerated SGD as the best optimiser substantially outperforming EM on large data sets.
+However, due to an incompatibility between Spark's programming paradigm and Sanchez' distributed SGD implementation, the performance of his program when using mini-batch optimisation in a distributed setting fails to meet expectations.
+
+Sanchez explains the context in which the dissertation is set very well in an iterative fashion.
+He first gives a high level overview of the dissertation and its topic in the introduction section.
+This includes a short description of the current state of affairs concerning mainstream GMM implementations, the problem statement, and the motivation of the dissertation project.
+He then goes on to elaborate on the technical background of the project in a dedicated section which constitutes a wider overview of related works.
+Finally, he presents the results of the previous work that the dissertation is directly based on in yet another dedicated section.
+
+The dissertation is fairly clear about the scope of the problem.
+It is explicitly stated in the introduction section and also explained in the results section.
+The approaches taken for the implementation of the program and its testing are also carefully described in the program design and results sections.
+The testing methods, featuring an automated framework and parameterised synthetic data generation, are remarkably sophisticated and comprehensive.
+The program is compared against a good standard in the form of the GMM implementation of Spark's ML library.
+Due to the fixed random seeds, the results are reproducible as well.
+However, Sanchez is not entirely clear about the metrics of success.
+Although the serial benchmarks show promising results, the distributed tests can probably be considered unsuccessful.
+Moreover, the data streaming performance test results are hard to interpret without additional context or a reference to compare against.
+
+The overall structure of the dissertation is well organised and intuitive.
+Sanchez does an excellent job explaining the technicalities in an easily understandable, straightforward manner.
+He manages this by using simple and concise sentences that build on top of each other and thus allow him to progressively convey complex information.
+Generally, the language he uses is objective and appropriate for a scientific report.
+With the exception of a few typos and grammar mistakes, his writing, punctuation, and overall quality of presentation are excellent.
+
+Sanchez' use of algorithms, tables, and figures is clean and fairly consistent.
+His figures are visually pleasing, well labeled, big enough to read the details, and most importantly, they aid understanding.
+Some of them could perhaps benefit from a grid and more visible axes, but the general quality of the figures is high.
+His algorithms and tables are all neatly formatted as well and all his captions are informative.
+Except for equation 25, which is numbered but never referenced, his use of equations throughout the report is clean and helpful.
+Finally, while Sanchez' references are well formatted and organised, some of his in-text citations are rather unconventional due to missing white spaces between the words in the text and the citations.
 

### 2020-04-17 report/MSc-report.tex
Update on Overleaf.

@@
 \usepackage{tabularx}
+\usepackage{amsmath}
 
@@
 
-\begin{equation}
+The first one of these milestones has been the implementation of a basic one dimensional diffusion simulation in C defined by the model
+\begin{equation*}
     \diffp{u}{t} = d \diffp{u}{{x^2}}
-\label{equation:1d_diff}
-\end{equation}
-
-The first one of these milestones has been the implementation of a basic one dimensional diffusion simulation in C defined by the model in Equation \ref{equation:1d_diff} where $u$ is the unknown function, $x$ is the spatial coordinate, $t$ is time, and $d$ is the diffusion coefficient.
+\end{equation*}
+where $u$ is the unknown function, $x$ is the spatial coordinate, $t$ is time, and $d$ is the diffusion coefficient.
 This simulation uses simple Dirichlet boundary conditions and Gaussian initial values.
@@
 
+The second, somewhat more ambitious milestone has been the implementation of a simple shared-memory Parareal framework in C using POSIX threads to solve an ordinary differential equation (ODE) describing the population growth of rabbits.
+This ODE is expressed by
 \begin{equation}
@@
 \end{equation}
-
-The second, somewhat more ambitious milestone has been the implementation of a simple shared-memory Parareal framework in C using POSIX threads to solve an ordinary differential equation (ODE) describing the population growth of rabbits.
-This ODE is expressed by Equation \ref{equation:rabbit_pop} where $R$ is the number of rabbits and $\alpha$ is the population growth rate.
+where $R$ is the number of rabbits and $\alpha$ is the population growth rate.
 The framework employs the forward Euler method as $G$ and RK4 as $F$ but it can only handle matching coarse and fine operator step sizes.
-Although the algorithm runs for a predefined number of corrective iterations, the implementation demonstrates a significant speedup at reasonable levels of accuracy even when running on only 4 threads.
+Although the algorithm runs for a predefined number of corrective iterations, the implementation demonstrates a significant speedup at reasonable levels of accuracy even when running on only four threads.
 Moreover, it showcases how the parallel solution converges to the serial fine solution as the number of iterations approaches $P$.
@@
 The ML operator trains the model on every new differential equation by generating a set of input-output pairs through a stochastic process using the target operator and fitting the regression model to the generated data.
-The data generation involves repeatedly integrating over the differential equation by the target operator and adding a $0$-mean Gaussian noise to the estimate at every time step to use as the initial value for the next step and thus perturbate the trajectory of the target operator's solution.
+The data generation involves repeatedly integrating over the differential equation by the target operator and adding a zero-mean Gaussian noise to the estimate at every time step to use as the initial value for the next step and thus perturbate the trajectory of the target operator's solution.
 Using this approach, we have successfully trained and utilised an ML operator backed by a linear regression model as the coarse operator $G$ in our Parareal framework to solve a simple ODE.
-The implementation of this framework and our experiment can be found in the \href{https://git.ecdf.ed.ac.uk/msc-19-20/s1984842/blob/master/code/python/}{\texttt{code/python}} directory of the project repository
+The implementation of this framework and our experiment can be found in the \href{https://git.ecdf.ed.ac.uk/msc-19-20/s1984842/blob/master/code/python/}{\texttt{code/python}} directory of the project repository.
+
+Table \ref{tab:coarse_times} shows the average execution times, based on ten runs, of two different coarse operators as they integrate over the entire time domain of the rabbit population problem defined by Equation \ref{equation:rabbit_pop} using the values $R_0 = 1000$, $\alpha = 0.0001$, and $t \in [0, 50000]$.
+The first operator is a conventional one using the explicit midpoint integration method and a step size of $0.25$ while the second operator is an ML accelerated one with a step size of $100$ backed by an sklearn \cite{pedregosa2011} linear regression model trained on data labelled by the first operator.
+As seen in the table, the ML accelerated operator solves the differential equation almost fourteen times faster with an average deviation of less than $10^{-6}$ from the conventional operator's solution.
+This performance difference is due to the fact that the ML operator's step size is $400$ times greater and therefore the solution is evaluated $400$ times fewer while still maintaining the same level of accuracy.
 
@@
 \midrule
-Run time (s) & 0.233 & \textbf{0.017}\\
+Run time / s & 0.233 & 0.017\\
 \bottomrule
@@
 
-Table \ref{tab:coarse_times} shows the average execution times, based on $10$ runs, of two different coarse operators as they integrate over the entire time domain of the rabbit population problem defined by Equation \ref{equation:rabbit_pop} using the values $R_0 = 1000$, $\alpha = 0.0001$, and $t \in [0, 50000]$.
-The first operator is a conventional one using the explicit midpoint integration method and a step size of $0.25$ while the second operator is an ML accelerated one with a step size of $100$ backed by an sklearn \cite{pedregosa2011} linear regression model trained on data labelled by the first operator.
-As seen in the table, the ML accelerated operator solves the differential equation almost $14$ times faster with an average deviation of less than $10^{-6}$ from the conventional operator's solution.
-This performance difference is due to the fact that the ML operator's step size is $400$ times greater and therefore the solution is evaluated $400$ times fewer while still maintaining the same level of accuracy.
+Table \ref{tab:parareal_times} shows the average execution times, again across ten runs, of different solvers on the same rabbit population problem.
+The first, serial solver is a conventional operator using the RK4 integration method and a step size of $0.01$.
+The others are all Parareal solvers running on four processes using the first solver as their fine operator and $0.001$ as their update threshold (which means that the algorithm finishes once the absolute value of the largest update to the solution at any point of the fine operator's temporal discretisation is below this threshold).
+The first Parareal solver uses the explicit midpoint operator from Table \ref{tab:coarse_times} as its coarse operator while the second and third Parareal solvers use the linear regression operator.
+Both the conventional and the ML accelerated Parareal solvers converge after just one corrective iteration and thus are able to achieve a significant speedup compared to the serial solver while producing a solution identical to the serial one to six and five significant figures respectively.
+While the ML accelerated operator is slightly less accurate, it still converges within the same number of iterations given our accuracy threshold and it actually solves the differential equation a whole second faster.
+In contrast, when taking the model training into account, the ML accelerated Parareal solver is barely  faster than the serial solver.
+However, the training time could potentially be significantly reduced for the given accuracy requirements by decreasing the volume of training data generated.
 
@@
 \midrule
-Run time (s) & 8.387 & 6.082 & \textbf{5.080} & 8.020\\
+Run time / s & 8.387 & 6.082 & 5.080 & 8.020\\
 \bottomrule
@@
 \end{table}
-
-Table \ref{tab:parareal_times} shows the average execution times, again across $10$ runs, of different solvers on the same rabbit population problem.
-The first, serial solver is a conventional operator using the RK4 integration method and a step size of $0.01$.
-The others are all Parareal solvers running on $4$ processes using the first solver as their fine operator and $0.001$ as their update threshold (which means that the algorithm finishes once the absolute value of the largest update to the solution at any point of the fine operator's temporal discretisation is below this threshold).
-The first Parareal solver uses the explicit midpoint operator from Table \ref{tab:coarse_times} as its coarse operator while the second and third Parareal solvers use the linear regression operator.
-Both the conventional and the ML accelerated Parareal solvers converge after just one corrective iteration and thus are able to achieve a significant speedup compared to the serial solver while producing a solution identical to the serial one to $6$ and $5$ significant figures respectively.
-While the ML accelerated operator is slightly less accurate, it still converges within the same number of iterations given our accuracy threshold and it actually solves the differential equation a whole second faster.
-In contrast, when taking the model training into account, the ML accelerated Parareal solver is barely  faster than the serial solver.
-However, the training time could potentially be significantly reduced for the given accuracy requirements by decreasing the volume of training data generated.
 
@@
 \begin{enumerate}
-    \item Implementation
+    \item \textbf{Implementation}:
         \begin{enumerate}
-        	\item Testing
-        	\item Systems of ODEs
-        	\item PDEs
-        	\item ML operators
-        	\item Performance optimisation
+        	\item \textbf{Testing}: This includes general software testing in the form of regression tests and unit tests of the components of our framework where possible.
+        	\item \textbf{Systems of ODEs}: The extension of our Parareal framework to handle systems of ODEs.
+This is synonymous with enabling differential equations to have vector valued unknown functions.
+As the first problem we intend to test our framework on, the Lotka-Volterra model, is a system of two ODEs, it is imperative that we complete this sub-task as early as possible.
+        	\item \textbf{PDEs}: The extension of our existing framework to handle PDEs.
+This requires differential equations to be able to specify boundary conditions and have multiple independent variables.
+The remaining two problems we plan to test our framework on are PDEs, therefore this is a high priority sub-task as well.
+        	\item \textbf{ML operators}: The implementation of the ML accelerated operators for the different model types and training processes (one for PINNs and one for all the other models).
+        	\item \textbf{Performance optimisation}: This includes hyperparameter tuning and any code optimisations to improve the performance of the operators and the Parareal framework.
         \end{enumerate}
-    \item Benchmarking
+    \item \textbf{Benchmarking}:
         \begin{enumerate}
-        	\item Only ML operators
-        	\item Framework with and without ML
-        	\item Against another parallel solver
+        	\item \textbf{Only ML operators}: The benchmarking of the different ML accelerated operators against conventional solvers on the three problems outlined in our project proposal.
+This includes the measurement of accuracy and training time as well.
+        	\item \textbf{Framework with and without ML}: The benchmarking of the framework using different ML accelerated coarse operators against the same framework using different conventional operators.
+As we intend to use fixed accuracy requirements, the primary metrics for this task are the wall time to solution and the model training time.
+        	\item \textbf{Against another parallel solver}: The comparison of the framework using different ML accelerated coarse operators against conventional parallel solvers using the same number of processes.
+The metrics for this task are the same as above with the addition of accuracy (as we most probably cannot specify accuracy requirements the same way between our Parareal implementation and other solvers).
         \end{enumerate}
-    \item Writing
+    \item \textbf{Writing}:
         \begin{enumerate}
-            \item Engineering log
-            \item Phase 1
-            \item Phase 2
-        	\item Proof reading
+            \item \textbf{Engineering log}: The maintenance of an engineering log during the implementation phase to keep track of design decisions and technical details.
+            \item \textbf{Phase 1}: The writing up of the problem formulation and the design and implementation sections of the dissertation.
+            \item \textbf{Phase 2}: The completion of the rest of the dissertation such as the introduction, background, results, and conclusions sections.
+        	\item \textbf{Proof reading}: The quality control of the dissertation report to make sure it is free of typos, grammar mistakes, and other errors.
         \end{enumerate}
@@
 
-For the implementation task, these sub-tasks are software testing, the extension of the Parareal framework to handle systems of ODEs and individual PDEs, the implementation of the two final ML accelerated operators discussed in the project proposal (one for PINNs and one for all the other models), and code optimisation along with hyperparameter tuning.
-As for the benchmarking task, the sub-tasks are the comparison of the ML accelerated operator to conventional operators, the comparison of the Parareal framework using an ML accelerated coarse operator to the same framework using a conventional coarse operator, and the comparison of the ML accelerated Parareal framework to other parallel solvers.
-Finally, the last main task, the writing of the dissertation consists of four smaller units of work.
-These are the engineering log to keep during the initial development period, the first phase for writing up the problem formulation and the design and implementation sections of the dissertation, the second phase for writing up the rest of the dissertation such as the literature review and the benchmark results, and lastly, the proof reading of the report.
+Figure \ref{fig:gantt} contains a Gantt chart displaying the timeline of the dissertation project and the different tasks it encompasses.
+Although the times are just rough and somewhat optimistic estimates, the chart sufficiently visualises the overlaps between certain tasks.
+For example, it clearly captures the non-linear nature of performance optimisation and benchmarking as they feed back into each other.
 
@@
 
-Figure \ref{fig:gantt} contains a Gantt chart displaying the timeline of the dissertation project and the different tasks it encompasses.
-Although the times are just rough and somewhat optimistic estimates, the chart sufficiently visualises the overlaps between certain tasks.
-For example, it clearly captures the non-linear nature of optimisation and benchmarking as they feed back into each other.
-
 As for hardware requirements, our plan is to use the author's personal computer for most of the implementation phase.
-This computer has $6$ physical cores which is sufficient for development and testing purposes.
+This computer has six physical cores which is sufficient for development and testing purposes.
 However for optimisation and benchmarking, we plan to use high performance computing services.
@@
 Additionally, if the new Cirrus GPUs are installed by late June, we also plan to take advantage of them for our neural network based models.
-For this, our required budget estimate is $50$ GPU hours.
+For this, our required budget estimate is $500$ GPU hours.
 
@@
 
-We identify the following as the main risks of the dissertation project.
-
-\begin{itemize}
-    \item Failure to implement a single framework to handle both ODEs and PDEs in time
-    \item Some of the model training approaches may not work on complex problems
-    \item Erratic model behaviour due to the stochastic nature of the training processes
-    \item Transfer learning may not work
-    \item Unavailability of computing resources
-    \item Disruptions due to Covid-19
-\end{itemize}
-
-The first risk we identify is the author's failure to implement a Parareal framework that can handle ODEs, systems of ODEs, and PDEs in time for our benchmarks that need to be performed early enough to leave sufficient time for writing up the results.
+The first risk of the dissertation project we identify is the author's failure to implement a Parareal framework that can handle ODEs, systems of ODEs, and PDEs in time for our benchmarks that need to be performed early enough to leave sufficient time for writing up the results.
 Failure to do this would likely restrict our benchmarks to ODEs only which are not representative of more complex real life problems.
@@
     		\item Machine learning accelerated operators
+    		\item Usage
     	\end{enumerate}
@@
 He then goes on to elaborate on the technical background of the project in a dedicated section which constitutes a wider overview of related works.
-Finally, he presents the results of the previous work that the dissertation is directly based on in yet another dedicated section.
+Finally, he presents the results of the previous work that his dissertation directly builds upon in yet another dedicated section.
+His approach is a good example for us to follow and is indeed what we base the outline of our dissertation report on.
 
-The dissertation is fairly clear about the scope of the problem.
+Sanchez' dissertation is fairly clear about the scope of the problem.
 It is explicitly stated in the introduction section and also explained in the results section.
@@
 Due to the fixed random seeds, the results are reproducible as well.
+These are all good practices that we intend to utilise for our dissertation, too.
 However, Sanchez is not entirely clear about the metrics of success.
